# ===== SCRIPT OVERVIEW =====
# Task:   Identify misinformation supersharers (users responsible for top 80% of
#         misinformation sharing) and compare their emotion features to other users.
#         PIPELINE NOTE: this script writes Austria-Panel-Supersharers.csv which is
#         required by pull_sources.py and pull_supersharer_tweets.py — run before them.
# Input:  data/Austria-Panel-Users-Daily.csv.gz
#         data/Austria-Panel-Users-Features-Tweet.csv
# Output: data/Austria-Panel-Supersharers.csv
#         ./users/supersharers_cdf.png
#         ./users/supersharers_emotion_comparison.png
#         output/figures/supersharers_cdf.pdf
#         output/figures/supersharers_emotion_comparison.pdf

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 6.5,
                      fig.height = 4.5)
set.seed(123)
```

```{r libraries}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(broom)
```

```{r}
custom_plot <- function() {
  theme_classic() +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x  = element_text(size = 11),
      axis.text.y  = element_text(size = 11),
      legend.title = element_text(size = 12),
      legend.text  = element_text(size = 11)
    )
}
```

```{r}
src <- "../../data"
dst <- "../../output"

dir.create("./users", showWarnings = FALSE, recursive = TRUE)
```

## User data
```{r}
users <- read_csv(file.path(src, "Austria-Panel-Users-Daily.csv.gz"),
                     col_types = cols(
                        author_id = col_character(),
                        day = col_character(),
                        .default = col_double())
                      ) 
nrow(users)
```

```{r}
features <- read_csv(file.path(src, "Austria-Panel-Users-Features-Tweet.csv"),
                     col_types = cols(
                        author_id = col_character(),
                        .default = col_double())
                      )

features$shares_news <- as.integer(features$n_news_shared > 0)
features$shares_biased <- as.integer(features$n_biased_shared > 0)
features$shares_untrustworthy <- as.integer(features$n_untrustworthy_shared > 0)
```

# ===== CONCENTRATION ANALYSIS =====
```{r}
# get total number of misinfo shared per user per day
users_summary <- users %>%
  group_by(author_id) %>%
  summarise(
    total_misinfo_count = sum(n_untrustworthy, na.rm = TRUE),
    total_trustworthy_count = sum(n_trustworthy, na.rm = TRUE),
    total_news_count = sum(n_news, na.rm = TRUE),
    total_biased_count = sum(n_biased, na.rm = TRUE)
  ) %>%
  ungroup()
  
head(users_summary,2)
```

```{r}
untrustworthy_by_user <- users_summary %>%
  dplyr:::select(author_id, total_misinfo_count) %>%  
  arrange(desc(total_misinfo_count)) %>%
  mutate(
    cumsum_count = cumsum(total_misinfo_count),
    cumsum_prop = cumsum_count / sum(total_misinfo_count),
    user_rank = row_number(),
    percentile = user_rank / n() * 100,
    source_type = "Untrustworthy"
  )

threshold_untrustworthy <- min(which(untrustworthy_by_user$cumsum_prop >= 0.80))
percentile_untrustworthy <- untrustworthy_by_user$percentile[threshold_untrustworthy]
untrustworthy_supersharers <- untrustworthy_by_user %>%
  slice(1:threshold_untrustworthy)
```

```{r}
trustworthy_by_user <- users_summary %>%
  dplyr:::select(author_id, total_trustworthy_count) %>%  
  filter(total_trustworthy_count > 0) %>%
  arrange(desc(total_trustworthy_count)) %>%
  mutate(
    cumsum_count = cumsum(total_trustworthy_count),
    cumsum_prop = cumsum_count / sum(total_trustworthy_count),
    user_rank = row_number(),
    percentile = user_rank / n() * 100,
    source_type = "Trustworthy"
  )

threshold_trustworthy <- min(which(trustworthy_by_user$cumsum_prop >= 0.80))
percentile_trustworthy <- trustworthy_by_user$percentile[threshold_trustworthy]

trustworthy_supersharers <- trustworthy_by_user %>%
  slice(1:threshold_trustworthy)
```


```{r}
users_combined <- bind_rows(untrustworthy_by_user, trustworthy_by_user)

# plot cumulative distribution
p_cd <- ggplot(users_combined, 
                aes(x = percentile, 
                    y = cumsum_prop,
                    color = source_type)) +
  geom_line(linewidth=1) +
  geom_hline(yintercept = 0.8, 
              linetype = "dashed", 
              color = "black") +
  scale_color_manual(
    values = c("Trustworthy" = "#2a9d8f", 
               "Untrustworthy" = "#8b7ab8")
  ) +
  labs(
    x = "users (%)",
    y = "news shared (%)",
    color = NULL
  ) +
  custom_plot()+
  # adjust legend position to be bottom right corner
  theme(legend.position = c(0.85, 0.1))+
  #annotation with number of users sharing 80% of misinfo
  annotate("text", 
           x = 15, y = 0.83,
           label = paste0(round(percentile_untrustworthy, 2), "% vs ", round(percentile_trustworthy, 2), "% of users"),
           size = 4,
           hjust = 0)

ggsave(file.path(dst, "supersharers_cdf.pdf"),
       plot = p_cd,
       width = 6.5, height = 4)
```


# ===== SUPERSHARER ANALYSIS =====

```{r}
# get the ppl sharing 80% of misinfo
# and add their features
untrustworthy_supersharers <- untrustworthy_supersharers %>%
  inner_join(features, by = "author_id") %>%
  mutate(group = "Untrustworthy")

# get people who are the supersharers of trustworthy info
trustworthy_supersharers <- trustworthy_supersharers %>%
  inner_join(features, by = "author_id") %>%
  mutate(group = "Trustworthy")

supersharers <- bind_rows(untrustworthy_supersharers, trustworthy_supersharers)
nrow(supersharers)
```
```{r}
# save as csv 
write_csv(supersharers, 
          file.path(src, "Austria-Panel-Supersharers.csv"))
```
```{r}
anger_data <- supersharers %>%
  dplyr:::select(author_id, group,
         anger_baseline, anger_variability, anger_instability, anger_inertia) %>%
  pivot_longer(cols = c(anger_baseline, anger_variability, anger_instability, anger_inertia),
               names_to = "metric",
               values_to = "value") %>%
  mutate(
    metric = gsub("anger_", "", metric),
    metric = factor(metric, levels = c("baseline", "variability", "instability", "inertia")),
    emotion = "Anger"
  )

fear_data <- supersharers %>%
  dplyr:::select(author_id, group,
         fear_baseline, fear_variability, fear_instability, fear_inertia) %>%
  pivot_longer(cols = c(fear_baseline, fear_variability, fear_instability, fear_inertia),
               names_to = "metric",
               values_to = "value") %>%
  mutate(
    metric = gsub("fear_", "", metric),
    metric = factor(metric, levels = c("baseline", "variability", "instability", "inertia")),
    emotion = "Fear"
  )

anger_stats <- anger_data %>%
  group_by(group, metric) %>%
  summarise(
    n = n(),
    mean_value = mean(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    metric_label = tools::toTitleCase(as.character(metric))
  )

# Calculate stats for fear
fear_stats <- fear_data %>%
  group_by(group, metric) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    std_value = sd(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    metric_label = tools::toTitleCase(as.character(metric))
  )

anger_plot_data <- anger_data %>%
  left_join(anger_stats, by = c("group", "metric"))

fear_plot_data <- fear_data %>%
  left_join(fear_stats, by = c("group", "metric"))
```


```{r}
run_tests <- function(data, emotion_name) {
  results <- data %>%
    group_by(metric_label) %>%
    nest() %>%
    mutate(
      # Wilcoxon test (non-parametric, good for distributions)
      wilcox_test = map(data, ~{
        trustworthy <- .x %>% filter(group == "Trustworthy") %>% pull(value)
        untrustworthy <- .x %>% filter(group == "Untrustworthy") %>% pull(value)
        wilcox.test(trustworthy, untrustworthy)
      }),
    
      p_value = map_dbl(wilcox_test, ~.x$p.value),
      significance = case_when(
        p_value < 0.001 ~ "***",
        p_value < 0.01 ~ "**",
        p_value < 0.05 ~ "*",
        TRUE ~ "ns"
      ),
      
      emotion = emotion_name
    ) %>%
    dplyr:::select(metric_label, p_value, significance, emotion)
  
  return(results)
}

anger_tests <- run_tests(anger_plot_data, "anger")
fear_tests <- run_tests(fear_plot_data, "fear")

all_tests <- bind_rows(anger_tests, fear_tests)

print(all_tests)
```


```{r}
# Calculate density and extract peaks for anger
anger_density <- anger_plot_data %>%
  group_by(metric_label, group) %>%
  do({
    d <- density(.$value, na.rm = TRUE)
    data.frame(x = d$x, density = d$y)
  }) %>%
  ungroup()

anger_peaks <- anger_density %>%
  group_by(metric_label, group) %>%
  slice_max(density, n = 1) %>%
  ungroup() %>%
  select(metric_label, group, peak_x = x, peak_y = density)

# Calculate density and extract peaks for fear
fear_density <- fear_plot_data %>%
  group_by(metric_label, group) %>%
  do({
    d <- density(.$value, na.rm = TRUE)
    data.frame(x = d$x, density = d$y)
  }) %>%
  ungroup()

fear_peaks <- fear_density %>%
  group_by(metric_label, group) %>%
  slice_max(density, n = 1) %>%
  ungroup() %>%
  select(metric_label, group, peak_x = x, peak_y = density)
```
```{r}
group_colors <- c(
  "Untrustworthy" = "#8b7ab8",
  "Trustworthy" = "#1b9e77"
)

group_sizes <- supersharers %>%
  group_by(group) %>%
  summarize(n = n(), .groups = "drop")

legend_labels <- group_sizes %>%
  mutate(legend_label = sprintf("%s (n=%s)", group, scales::comma(n)))

legend_mapping <- setNames(legend_labels$legend_label, group_sizes$group)

anger_stars <- anger_tests %>%
  left_join(
    anger_stats %>% 
      select(metric_label, group, mean_value) %>%
      pivot_wider(names_from = group, values_from = mean_value),
    by = "metric_label"
  ) %>%
  left_join(
    anger_peaks %>%
      group_by(metric_label) %>%
      slice_max(peak_y, n = 1) %>%
      dplyr:::select(metric_label, peak_y),
    by = "metric_label"
  ) %>%
  mutate(
    x = pmax(Trustworthy, Untrustworthy) + 0.01,  
    y = peak_y * 0.99, 
    label = significance
  )

fear_stars <- fear_tests %>%
  left_join(
    fear_stats %>% 
      select(metric_label, group, mean_value) %>%
      pivot_wider(names_from = group, values_from = mean_value),
    by = "metric_label"
  ) %>%
  left_join(
    fear_peaks %>%
      group_by(metric_label) %>%
      slice_max(peak_y, n = 1) %>%
      select(metric_label, peak_y),
    by = "metric_label"
  ) %>%
  mutate(
    x = pmax(Trustworthy, Untrustworthy) + 0.01,
    y = peak_y * 0.99,
    label = significance
  )

p_anger <- ggplot(anger_plot_data, aes(x = value, fill = group, color = group)) +
  geom_density(alpha = 0.4, linewidth = 0.5) +
  geom_vline(data = anger_stats,
             aes(xintercept = mean_value, 
             color = group), 
             linetype = "dashed", linewidth = 0.8) +
  geom_text(data = anger_stats %>% filter(group == "Untrustworthy"),
            aes(x = Inf, y = Inf, 
                label = paste0("M=", round(mean_value, 3))),
            vjust = 1.3, hjust = 2.2, size = 3, 
            color = group_colors["Untrustworthy"],
            inherit.aes = FALSE) +
  geom_text(data = anger_stats %>% filter(group == "Trustworthy"),
            aes(x = Inf, y = Inf, 
                label = paste0("M=", round(mean_value, 3))),
            vjust = 2.9, hjust = 2.2, size = 3, 
            color = group_colors["Trustworthy"],
            inherit.aes = FALSE) +
  geom_text(data = anger_stars,
            aes(x = x, y = y, label = label),
            size = 3, 
            hjust = 0,
            vjust = 0.5,
            inherit.aes = FALSE) +
  facet_wrap(~metric_label, ncol = 1, 
                            scales = "free_y", 
                            strip.position = "left") +
  scale_x_continuous(limits = c(-0.25, 0.75)) +
  scale_fill_manual(values = group_colors, labels = legend_mapping) +  # Added labels
  scale_color_manual(values = group_colors, labels = legend_mapping) + 
  labs(x = "Value", y = NULL, 
       fill = NULL, 
       color = NULL, 
       title = "a. anger") +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
   strip.text.y.left = element_text(size = 8, angle = 90, 
                                    hjust = 0.5, vjust = 0.5),
    strip.background = element_rect(
                          fill = "white", 
                          color = "black"),
    legend.position = c(1.1, -0.14),  # x=0.5 (center), y=1.02 (slightly above plot)
    legend.direction = "horizontal",
    plot.margin = margin(t = 5, r = 5, b = 20, l = 5, unit = "pt"),
    axis.text.x = element_text(size = 7),
    axis.line.x = element_line(color = "black"),
    panel.spacing = unit(0.5, "lines"),
    plot.title = element_text(size = 11, face = "bold")
  )

# Create fear plot (Plot B)
p_fear <- ggplot(fear_plot_data,
                  aes(x = value, 
                      fill = group, 
                      color = group)) +
  geom_density(alpha = 0.4, linewidth = 0.5) +
  geom_vline(data = fear_stats,
             aes(xintercept = mean_value,
                 color = group), 
                 linetype = "dashed", 
                 linewidth = 0.8) +
  geom_text(data = fear_stats %>% filter(group == "Untrustworthy"),
            aes(x = Inf, y = Inf, 
                label = paste0("M=", round(mean_value, 3))),
            vjust = 1.3, hjust = 2.2, size = 3, color = group_colors["Untrustworthy"],
            inherit.aes = FALSE) +
  geom_text(data = fear_stats %>% filter(group == "Trustworthy"),
            aes(x = Inf, y = Inf, 
                label = paste0("M=", round(mean_value, 3))),
            vjust = 2.9, hjust = 2.2, size = 3, color = group_colors["Trustworthy"],
            inherit.aes = FALSE) +
  geom_text(data = fear_stars,
            aes(x = x, y = y, label = label),
            size = 3, 
            hjust = 0,
            vjust = 0.5,
            inherit.aes = FALSE) +
  facet_wrap(~metric_label, 
                ncol = 1, 
                scales = "free_y", 
                strip.position = "left") +
  scale_x_continuous(limits = c(-0.25, 0.75)) +
  scale_fill_manual(values = group_colors) +
  scale_color_manual(values = group_colors) +
  labs(x = "Value", y = NULL, 
       fill = "Group", 
       color = "Group",
       title = "b. fear") +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    strip.text.y.left = element_text(size = 8, angle = 90, hjust = 0.5, vjust = 0.5),
    strip.background = element_rect(fill = "white", color = "black"),
    legend.position = "none",
    axis.text.x = element_text(size = 7),
    axis.line.x = element_line(color = "black"),
    panel.spacing = unit(0.5, "lines"),
    plot.title = element_text(size = 11, face = "bold")
  )


p_combined <- p_anger + p_fear + plot_layout(ncol = 2)
print(p_combined)
```

```{r}
# save as pdf in paper repository
ggsave(file.path(dst, "supersharers_emotion_comparison.pdf"),
       plot = p_combined,
       width = 6.5, height = 5)
```

```{r}
# run all above
```
