# ===== SCRIPT OVERVIEW =====
# Task:   Analyse the reply network structure: summary statistics, backbone visualisation,
#         and separate networks for untrustworthy vs. trustworthy news supersharers.
#         Produces Figure S18 and Table S23 (network statistics comparison).
# Input:  data/Austria-Panel-Network-Edges.csv.gz       (output of 12a-pull_replies.py)
#         data/Austria-Panel-Network-Nodes.csv.gz       (output of 12a-pull_replies.py)
#         data/Austria-Panel-Network-Edges-Internal.csv (output of 12b-pull_replies_internal.py)
#         data/Austria-Panel-Network-Nodes-Internal.csv (output of 12b-pull_replies_internal.py)
#         data/Austria-Panel-Supersharers.csv           (output of 11-analyze_supersharers.Rmd)
#         data/users/reply_analysis_summary.csv         (output of 12a-pull_replies.py)
# Output: output/figures/edge_node_distributions.pdf
#         output/figures/combined_supersharer_networks.png
#         output/tables/network_comparison.tex
#         output/tables/replies_latex.tex

```{r setup}
options(scipen=999)
library(kableExtra)
library(tidyverse)
library(igraph)
library(ggraph)
library(patchwork)

set.seed(636)
```

```{r paths}
src <- "../../data"
dst <- "../../output"
```

```{r custom_plot}
custom_plot <- function(plot) {
  theme_classic() +
    theme(
      axis.title.x = element_text(size = 9),
      axis.title.y = element_text(size = 9),
      axis.text.x  = element_text(size = 8),
      axis.text.y  = element_text(size = 8),
      legend.title = element_text(size = 8),
      legend.text  = element_text(size = 8)
    )
}
```

```{r load_summary}
reply_summary <- read.csv(file.path(src, "users", "reply_analysis_summary.csv"))
head(reply_summary)
```

## Summary of reply analysis
```{r summary_table}
reply_summary <- reply_summary %>%
  mutate(Metric = case_when(
    Metric == "Unique users replied to" ~ "Users replied to",
    Metric == "Users replied to IN sample" ~ "Users replied to in sample",
    Metric == "Users replied to OUTSIDE sample" ~ "Users replied to outside sample",
    Metric == "Overlap percentage" ~ "Overlap percentage",
    Metric == "Replies to users IN sample" ~ "Replies to users in sample",
    Metric == "Replies to users OUTSIDE sample" ~ "Replies to users outside sample",
    TRUE ~ Metric
  ))

# Table S24: Descriptive statistics for Austrian Twitter sample reply network
replies_latex <- kable(reply_summary,
                    format = "latex",
                    booktabs = TRUE,
                    align = c("l", "c", "r"),
                    escape = TRUE)

writeLines(as.character(replies_latex), file.path(dst,"replies_latex.tex"))
```

## Network summary of all users

```{r load_network}
edges <- read_csv(file.path(src, "Austria-Panel-Network-Edges.csv.gz"))
nodes <- read_csv(file.path(src, "Austria-Panel-Network-Nodes.csv.gz"))
nrow(edges); nrow(nodes)
```

```{r plot_network_summary}
p_edges <- ggplot(edges, aes(x = edge_type, fill = edge_type)) +
  geom_bar() +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "# of edges", x = "reply direction") +
  scale_fill_brewer(palette = "Set2") +
  custom_plot() +
  theme(legend.position = "none")

p_nodes <- ggplot(nodes, aes(x = n_replies_received + 1, fill = sample_status)) +
  geom_histogram(bins = 50) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_log10(labels = scales::comma) +
  facet_wrap(~sample_status) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "# of replies received (log scale)", y = "# of users") +
  custom_plot() +
  theme(legend.position = "none")

p_combined <- p_edges + p_nodes + plot_layout(widths = c(1, 2))

ggsave(file.path(dst, "edge_node_distributions.pdf"), p_combined,
       width = 6.5, height = 3.5)
```

## Separate networks for supersharers of untrustworthy vs. trustworthy news
```{r load_internal}
i_edges <- read_csv(file.path(src, "Austria-Panel-Network-Edges-Internal.csv"))
i_nodes <- read_csv(file.path(src, "Austria-Panel-Network-Nodes-Internal.csv"))
nrow(i_edges); nrow(i_nodes)
```

```{r load_supersharers}
supersharers <- read_csv(file.path(src, "Austria-Panel-Supersharers.csv"),
                       col_types = cols(
                         author_id = col_double(),
                         group = col_factor(levels = c("Trustworthy", "Untrustworthy"))
                       ))

untrustworthy_sharers <- supersharers %>%
  filter(group == "Untrustworthy") %>%
  pull(author_id)

trustworthy_sharers <- supersharers %>%
  filter(group == "Trustworthy") %>%
  pull(author_id)

untrustworthy_edges <- i_edges %>%
  filter(author_id %in% untrustworthy_sharers &
         in_reply_to_user_id %in% untrustworthy_sharers)

untrustworthy_nodes <- i_nodes %>%
  filter(user_id %in% c(untrustworthy_edges$author_id,
                        untrustworthy_edges$in_reply_to_user_id))

nrow(untrustworthy_edges); nrow(untrustworthy_nodes)
```

```{r plot_untrustworthy_network}
g_untrustworthy <- graph_from_data_frame(
  d = untrustworthy_edges,
  vertices = untrustworthy_nodes,
  directed = TRUE
)

p_untrustworthy <- ggraph(g_untrustworthy, layout = "fr") +
  geom_edge_link(aes(width = log(weight)), alpha = 0.6, color = "gray70") +
  geom_node_point(aes(size = log(total_degree + 1)), color = "#8b7ab8", alpha = 0.8) +
  scale_edge_width_continuous(range = c(0.2, 2)) +
  scale_size_continuous(range = c(0.5, 5)) +
  labs(title = "a. supersharers of 80%\n untrustworthy vs.",
       subtitle = sprintf("(%s edges, %s nodes)",
                         nrow(untrustworthy_edges), nrow(untrustworthy_nodes))) +
  theme_graph() +
  theme(legend.position = "none")
```

```{r plot_trustworthy_network}
trustworthy_edges <- i_edges %>%
  filter(author_id %in% trustworthy_sharers &
         in_reply_to_user_id %in% trustworthy_sharers)

trustworthy_nodes <- i_nodes %>%
  filter(user_id %in% c(trustworthy_edges$author_id,
                        trustworthy_edges$in_reply_to_user_id))
nrow(trustworthy_edges); nrow(trustworthy_nodes)

g_trustworthy <- graph_from_data_frame(
  d = trustworthy_edges,
  vertices = trustworthy_nodes,
  directed = TRUE
)

p_trustworthy <- ggraph(g_trustworthy, layout = "fr") +
  geom_edge_link(aes(width = log(weight)), alpha = 0.5, color = "gray70") +
  geom_node_point(aes(size = log(total_degree + 1)), color = "#1b9e77", alpha = 0.8) +
  scale_edge_width_continuous(range = c(0.2, 2)) +
  scale_size_continuous(range = c(0.5, 5)) +
  labs(title = "b. 80% of trustworthy news",
       subtitle = sprintf("(%s edges, %s nodes)",
                         nrow(trustworthy_edges), nrow(trustworthy_nodes))) +
  theme_graph() +
  theme(legend.position = "none")

p_combined_networks <- p_untrustworthy + p_trustworthy +
  plot_layout(widths = c(1, 2), ncol = 2)

ggsave(file.path(dst, "combined_supersharer_networks.png"),
       p_combined_networks, width = 6.5, height = 4)
```

### Network statistics comparison (Table S23)
```{r network_stats}
compute_network_stats <- function(g, network_name) {
  n_nodes <- vcount(g)
  n_edges <- ecount(g)
  density <- edge_density(g)
  deg_all <- degree(g, mode = "all")
  reciprocity_val <- reciprocity(g)
  components_weak <- components(g, mode = "weak")
  gcc_size <- max(components_weak$csize)
  gcc_prop <- gcc_size / n_nodes
  gcc_nodes <- which(components_weak$membership == which.max(components_weak$csize))
  g_gcc <- induced_subgraph(g, gcc_nodes)
  transitivity_global <- transitivity(g, type = "global")
  avg_path_length <- if (vcount(g_gcc) > 0 && ecount(g_gcc) > 0)
    mean_distance(g_gcc, directed = TRUE) else NA

  data.frame(
    Network = network_name,
    Nodes = n_nodes,
    Edges = n_edges,
    Density = density,
    Mean_Degree = mean(deg_all),
    Reciprocity = reciprocity_val,
    Giant_Component_Size = gcc_size,
    Giant_Component_Proportion = gcc_prop,
    Global_Clustering = transitivity_global,
    Avg_Path_Length = avg_path_length,
    Max_Kcore = max(coreness(g, mode = "all"))
  )
}

untrustworthy_stats <- compute_network_stats(g_untrustworthy, "Untrustworthy")
trustworthy_stats <- compute_network_stats(g_trustworthy, "Trustworthy")

network_comparison <- rbind(untrustworthy_stats, trustworthy_stats)
print(network_comparison)

summary_table <- data.frame(
  Metric = c("Nodes", "Edges", "Density", "Mean Degree",
             "Reciprocity", "Giant Component (%)",
             "Global Clustering", "Avg Path Length", "Max K-core"),
  Untrustworthy = c(
    format(untrustworthy_stats$Nodes, big.mark = ","),
    format(untrustworthy_stats$Edges, big.mark = ","),
    sprintf("%.3f", untrustworthy_stats$Density),
    sprintf("%.3f", untrustworthy_stats$Mean_Degree),
    sprintf("%.3f", untrustworthy_stats$Reciprocity),
    sprintf("%.1f%%", untrustworthy_stats$Giant_Component_Proportion * 100),
    sprintf("%.3f", untrustworthy_stats$Global_Clustering),
    sprintf("%.3f", untrustworthy_stats$Avg_Path_Length),
    untrustworthy_stats$Max_Kcore
  ),
  Trustworthy = c(
    format(trustworthy_stats$Nodes, big.mark = ","),
    format(trustworthy_stats$Edges, big.mark = ","),
    sprintf("%.3f", trustworthy_stats$Density),
    sprintf("%.3f", trustworthy_stats$Mean_Degree),
    sprintf("%.3f", trustworthy_stats$Reciprocity),
    sprintf("%.1f%%", trustworthy_stats$Giant_Component_Proportion * 100),
    sprintf("%.3f", trustworthy_stats$Global_Clustering),
    sprintf("%.3f", trustworthy_stats$Avg_Path_Length),
    trustworthy_stats$Max_Kcore
  )
)

# Table S23: Network statistics comparison (untrustworthy vs. trustworthy supersharers)
summary_latex <- kable(summary_table,
                      format = "latex",
                      booktabs = TRUE,
                      align = c("l", "r", "r"),
                      escape = TRUE)

writeLines(as.character(summary_latex),
           file.path(dst,  "network_comparison.tex"))
```
