---
title: "User Cohort Analysis"
output: html_document
---

# ===== SCRIPT OVERVIEW =====
# Task:   Analyse whether the 2022 structural breaks in untrustworthy news sharing
#         are driven by new accounts joining vs. behaviour change in existing users.
#         Produces Figure S10.
# Input:  data/Austria-Panel-User-Statistics.csv  (output of 10-aggregate_user_stats.py)
#         data/Austria-Panel-Users-Daily.csv.gz   (output of 4-extract_features.py)
# Output: output/figures/user_joins.pdf  (Figure S10)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(lubridate)
library(zoo)
library(patchwork)
library(scales)

set.seed(636)
```

```{r paths}
src <- "../../data"
dst <- "../../output"
```

```{r custom_plot}
custom_plot <- function() {
  theme_classic() +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x  = element_text(size = 11),
      axis.text.y  = element_text(size = 11),
      legend.title = element_text(size = 12),
      legend.text  = element_text(size = 11)
    )
}
```

```{r load_data}
user_data <- read_csv(file.path(src, "Austria-Panel-User-Statistics.csv"))

daily_data <- read_csv(file.path(src, "Austria-Panel-Users-Daily.csv.gz"),
                      col_types = cols(
                        author_id = col_double(),
                        day = col_character(),
                        .default = col_double()
                      )) %>%
  mutate(day = as.Date(day))
```

# Have new users joined over time?

```{r plot_joins}
user_data <- user_data %>%
  mutate(
    account_created_at = as.Date(account_created),
    first_tweet_date = as.Date(first_tweet_date),
    user_type = case_when(
      account_created_at >= as.Date("2022-01-01") ~ "new",
      TRUE ~ "old"
    )
  )

user_joins <- user_data %>%
  mutate(join_month = floor_date(account_created_at, "month")) %>%
  group_by(join_month) %>%
  summarise(n_new_users = n()) %>%
  filter(join_month >= as.Date("2019-01-01"))

p_joins <- ggplot(user_joins, aes(x = join_month, y = n_new_users)) +
  geom_col(fill = "steelblue") +
  geom_vline(xintercept = as.Date("2022-07-01"),
             linetype = "dashed", color = "red") +
  geom_vline(xintercept = as.Date("2022-08-28"),
             linetype = "dashed", color = "red") +
  labs(title = "a", x = "month", y = "# of users") +
  custom_plot() +
  theme(plot.title = element_text(size = 16, face = "bold"))

post_change_users <- daily_data %>%
  filter(day >= as.Date("2022-07-01")) %>%
  distinct(author_id) %>%
  left_join(user_data %>% select(author_id, account_created_at), by = "author_id") %>%
  mutate(
    account_age_at_july2022 = as.numeric(as.Date("2022-07-01") - account_created_at) / 365
  )

p_ages <- ggplot(post_change_users %>%
            filter(!is.na(account_age_at_july2022) & account_age_at_july2022 >= 0),
             aes(x = account_age_at_july2022)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  labs(title = "b", x = "account age (yrs) as of July 1, 2022", y = "# of users") +
  custom_plot() +
  theme(plot.title = element_text(size = 16, face = "bold"))
```

```{r plot_activity_by_type}
user_colors <- c("new" = "tomato", "old" = "steelblue")

activity_by_type <- daily_data %>%
  left_join(user_data %>% select(author_id, user_type), by = "author_id") %>%
  group_by(day, user_type) %>%
  summarise(
    avg_tweets = mean(n_tweets, na.rm = TRUE),
    avg_untrustworthy = mean(prop_untrustworthy, na.rm = TRUE),
    .groups = "drop"
  )

p_activity <- ggplot(activity_by_type %>% filter(day >= as.Date("2022-01-01")),
                     aes(x = day, y = avg_tweets, color = user_type)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = as.Date("2022-07-01"), linetype = "dashed", color = "red") +
  geom_vline(xintercept = as.Date("2022-08-28"), linetype = "dashed", color = "red") +
  scale_color_manual(values = user_colors, name = "account age",
                     labels = c("new" = "new", "old" = "old (>1 year)")) +
  labs(title = "c", x = "date", y = "# of tweets") +
  custom_plot() +
  theme(legend.position = "none",
        plot.title = element_text(size = 16, face = "bold"))

p_decomposition <- ggplot(activity_by_type %>% filter(day >= as.Date("2022-01-01")),
                          aes(x = day, y = avg_untrustworthy, color = user_type)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = as.Date("2022-07-01"), linetype = "dashed", color = "red") +
  geom_vline(xintercept = as.Date("2022-08-28"), linetype = "dashed", color = "red") +
  scale_color_manual(values = user_colors, name = "account age",
                     labels = c("new" = "new", "old" = "old (>1 year)")) +
  labs(title = "d", x = "date", y = "untrustworthy (%)") +
  custom_plot() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 16, face = "bold"))

ggsave(file.path(dst, "user_joins.pdf"),
       p_joins / p_ages / p_activity / p_decomposition,
       width = 6.5, height = 6.5, dpi = 300)
```
