# ===== SCRIPT OVERVIEW =====
# Task:   Analyse domain trustworthiness distribution and changes over time.
#         Focuses on June–September 2022 period.
# Input:  data/Austria-Panel-Tweets-Domains-Change-2022.csv.gz
#         data/Austria-Panel-Daily-Hashtags.csv
# Output: ./domains/domains_classification_prop.png
#         ./domains/domains_classification_raw.png / .pdf
#         ./domains/untrustworthy_domains_spikes.png
#         ./domains/top_hashtags_on_spike_days.csv
#         output/figures/domains_classification_prop.pdf
#         output/figures/untrustworthy_domains_spikes.pdf

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.width = 6.5, 
                      fig.height = 4.5)
options(scipen=999)
```

```{r load packages}
library(tidyverse)
library(dplyr)
library(patchwork)
library(kableExtra)
```

```{r paths}
src <- "../../data"
dst <- "../../output"
```

```{r custom plots}
custom_plot <- function() {
  theme_classic() +  
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x  = element_text(size = 11),
      axis.text.y  = element_text(size = 11),
      legend.title = element_text(size = 12),
      legend.text  = element_text(size = 11)
    )
}

```

```{r filter data}
# filter the main tweet-domain file to the Jun–Oct 2022 structural break period
authors_filtered <- read_csv(file.path(src, "authors_filtered.csv"),
                             col_types = cols(author_id = col_character())) %>%
  pull(author_id)

read_csv(
  file.path(src, "Austria-Panel-Tweets-Domains-PO-Dynamic.csv.gz"),
  col_select = c(author_id, created_at, ng_domain, Score),
  col_types = cols(author_id = col_character(),
                   created_at = col_character(),
                   ng_domain  = col_character(),
                   Score      = col_double())
) %>%
  filter(!is.na(ng_domain), !is.na(Score)) %>%
  filter(author_id %in% authors_filtered) %>%
  mutate(day = as.Date(str_sub(created_at, 1, 10))) %>%
  filter(day >= as.Date("2022-06-01"), day < as.Date("2022-11-01")) %>%
  select(day, ng_domain, Score) %>%
  write_csv(file.path(src, "Austria-Panel-Tweets-Domains-Change-2022.csv.gz"))
```

```{r}
domains <- read.csv(file.path(src, "Austria-Panel-Tweets-Domains-Change-2022.csv.gz"))

domains_per_day <- domains %>%
  separate_rows(ng_domain, sep = ", ") %>%
  group_by(day, ng_domain) %>%
  summarise(
    Score = first(Score),
    n_used = n(),
    .groups = 'drop')
```

```{r}
cutoff_july <- as.Date("2022-07-01")
cutoff_august <- as.Date("2022-08-28")

untrustworthy_trends <- domains_per_day %>%
  filter(Score < 60) %>%
  arrange(day, desc(n_used))

untrustworthy_before_after <- domains_per_day %>%
  filter(Score < 60) %>%
  mutate(period = if_else(day < cutoff_july, "before", "after")) %>%
  group_by(period, ng_domain, Score) %>%
  summarise(total_uses = sum(n_used), .groups = 'drop') %>%
  pivot_wider(names_from = period, values_from = total_uses, values_fill = 0) %>%
  mutate(
    change = after - before,
    pct_change = if_else(before > 0, (after - before) / before * 100, NA_real_)
  ) %>%
  arrange(desc(after))

untrustworthy_before_after%>%
  arrange(desc(pct_change))%>%
  print(n=30)
```


## Classify untrustworthy domains 
```{r}
# step 1: Get scores in the week before and week after cutoff for each domain
domain_classification <- domains_per_day %>%
  arrange(ng_domain, day) %>%
  group_by(ng_domain) %>%
  mutate(
    prev_score = lag(Score),
    first_appearance = (row_number() == 1),
    classification = case_when(
      # first time this domain appears and score < 60
      first_appearance & Score < 60 ~ "first",
      # previously had score >= 60, now < 60
      !is.na(prev_score) & prev_score >= 60 & Score < 60 ~ "downrated",
      # previously had score < 60, still < 60
      !is.na(prev_score) & prev_score < 60 & Score < 60 ~ "stable",
      TRUE ~ "Other"
    )
  ) %>%
  ungroup()
```

```{r}
# Step 2: Apply classification to ALL daily data where Score < 60
stacked_data_daily <- domain_classification %>%
  filter(Score < 60 & classification %in% c("first", "downrated", "stable")) %>%
  group_by(day, classification) %>%
  summarise(n_domains = sum(n_used), .groups = 'drop') %>%
  mutate(day = as.Date(day))

# Step 3: Plot proportions over ALL time
p_proportions <- ggplot(stacked_data_daily, aes(x = day, y = n_domains, fill = classification)) +
  geom_area(position = "fill") +
  scale_fill_manual(values = c(
    "first" = "#8da0cb", 
    "downrated" = "#fc8d62", 
    "stable" = "#66c2a5"
  )) +
  scale_y_continuous(labels = scales::percent) +
  geom_vline(xintercept = cutoff_july, linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = cutoff_august, linetype = "dashed", color = "black", size = 1) +
  labs(
    x = "time in days",
    y = "untrustworthy domains\nshared (%)",
    fill = "classification",
  ) +
  custom_plot()

ggsave(file.path(dst,"domains_classification_prop.pdf"), 
       plot = p_proportions, 
       width = 6.5, height = 3)
```

```{r}
# Step 4: Plot raw numbers
p_raw <- ggplot(stacked_data_daily, aes(x = day, y = n_domains, fill = classification)) +
  geom_area() +
  scale_fill_manual(values = c(
    "first" = "#8da0cb", 
    "downrated" = "#fc8d62", 
    "stable" = "#66c2a5"
  )) +
  labs(
    x = "time in days",
    y = "untrustworthy domains\nshared (#)",
    fill = "classification",
  ) +
  geom_vline(xintercept = cutoff_july, linetype = "dashed", color = "black", size = 1) +
  geom_vline(xintercept = cutoff_august, linetype = "dashed", color = "black", size = 1) +
  custom_plot()

#save as pdf
ggsave(file.path(dst, "domains_classification_raw.pdf"),
       plot = p_raw, 
       width = 6.5, height = 3)
```
```{r}
# Calculate total untrustworthy domains per day
daily_totals <- stacked_data_daily %>%
  group_by(day) %>%
  summarise(total_untrustworthy = sum(n_domains)) %>%
  arrange(day)

# Identify spike days - let's use 90th percentile as threshold
spike_threshold <- quantile(daily_totals$total_untrustworthy, 0.9, na.rm = TRUE)

spike_days <- daily_totals %>%
  filter(total_untrustworthy > spike_threshold) %>%
  pull(day)

print(spike_days)
```
```{r}
p_spikes <- ggplot(daily_totals, aes(x = day, y = total_untrustworthy)) +
  geom_line() +
  geom_hline(yintercept = spike_threshold, linetype = "dashed", color = "red") +
  geom_point(data = daily_totals %>% filter(day %in% spike_days), 
             aes(x = day, y = total_untrustworthy), color = "red", size = 2) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  labs(x = "time", y = "untrustworthy domains\n shared (#)") +
  custom_plot()

ggsave(file.path(dst, "untrustworthy_domains_spikes.pdf"),
         plot = p_spikes, 
         width = 6.5, height = 3)
```


## Domains on spike days
```{r}
# which untrustworthy domains were shared on the spike days? 
domains_on_spike_days <- domains_per_day %>%
  mutate(day = as.Date(day)) %>%
  filter(day %in% spike_days & Score < 60) %>%
  group_by(ng_domain, Score) %>%
  summarise(
    total_uses = sum(n_used),
    n_spike_days = n_distinct(day),
    avg_uses_per_spike_day = mean(n_used),
    .groups = 'drop'
  ) %>%
  arrange(desc(n_spike_days))

print(domains_on_spike_days, n = 10)
```

```{r}
# compare to baseline - what domains are shared more on spike days vs normal days?
normal_days <- daily_totals %>%
  filter(total_untrustworthy <= spike_threshold) %>%
  pull(day)

domains_normal_days <- domains_per_day %>%
  mutate(day = as.Date(day)) %>%
  filter(day %in% normal_days & Score < 60) %>%
  group_by(ng_domain) %>%
  summarise(
    avg_uses_normal = mean(n_used),
    .groups = 'drop'
  )

comparison <- domains_on_spike_days %>%
  left_join(domains_normal_days, by = "ng_domain") %>%
  mutate(
    spike_increase = avg_uses_per_spike_day - coalesce(avg_uses_normal, 0),
    spike_ratio = avg_uses_per_spike_day / coalesce(avg_uses_normal, 0.1)
  ) %>%
  arrange(desc(spike_increase))

# print the new columns
print(comparison %>% dplyr::select(ng_domain, avg_uses_per_spike_day, avg_uses_normal, spike_increase, spike_ratio), n = 10)
```


## Check out hashtags on those days
```{r}
hashtags_df <- read.csv(file.path(src, "Austria-Panel-Daily-Hashtags.csv")) %>%
  mutate(day = as.Date(day))%>%
  filter(day %in% spike_days)
```

```{r}
top_3_per_day <- hashtags_df %>%
  mutate(hashtag_list = str_split(hashtags, ",")) %>%
  rowwise() %>%
  mutate(
    top_3 = list(
      hashtag_list %>%
        table() %>%
        sort(decreasing = TRUE) %>%
        names() %>%
        head(3)
    )
  ) %>%
  ungroup() %>%
  mutate(top_3 = map(top_3, ~ if(length(.x) < 3) c(.x, rep(NA, 3 - length(.x))) 
                              else .x)) %>%
  unnest_wider(top_3, names_sep = "_") %>%
  rename(top_1 = top_3_1, top_2 = top_3_2, top_3 = top_3_3)

top_3_per_day %>%
  dplyr::select(day, top_1, top_2, top_3) %>%
  write.csv(file = file.path("domains", "top_hashtags_on_spike_days.csv"),
            row.names = FALSE)
```


