# ===== SCRIPT OVERVIEW =====
# Task:   Describe and summarise the univariate time series; compute proportions
#         and test for stationarity. Creates prop_bias = (count_right + count_left) / tweet_count.
#         Detects outliers in tweet count and saves them for downstream hashtag analysis.
# Input:  DATA_DIR/Austria-Panel-Daily.csv
# Output: OUTPUT_DIR/score_intervention.pdf
#         DATA_DIR/Austria-Panel-Count-Outliers.csv  (input for 16b-analyze_hashtags.Rmd)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 6, 
                      fig.height = 3,
                      dpi=300)

options(scipen=999)
```

```{r load packages}
library(readr)
library(scales)
library(tidyverse)
library(dplyr)
library(patchwork)
library(forecast)
library(tseries)
library(TSA)
library(lmtest)
library(tsoutliers)
library(fracdiff)
library(urca)
library(kableExtra)
```

```{r paths}
src <- "../../data"
dst <- "../../output"
dir.create(dst, recursive = TRUE, showWarnings = FALSE)
```

```{r custom plots}
custom_plot <- function() {
  theme_classic() +  
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x  = element_text(size = 11),
      axis.text.y  = element_text(size = 11),
      legend.title = element_text(size = 12),
      legend.text  = element_text(size = 11)
    )
}
```

# Exploration of time series data

```{r}
daily_summary <- read.csv(file.path(src,"Austria-Panel-Daily.csv"))
head(daily_summary, 2)
```

```{r}
daily_summary$day <- as.Date(daily_summary$day, format = "%Y-%m-%d")
summary(daily_summary)
```

```{r}
daily_summary <- daily_summary %>% arrange(day)
head(daily_summary, 5)
```

```{r}
start_date <- min(daily_summary$day)
start_year <- year(start_date)
start_day  <- yday(start_date)  # day of year for start
```

```{r}
# create proportions
daily_summary <- daily_summary %>%
  mutate(
    prop_news = (count_ng_domains / tweet_count) * 100,
    prop_misinfo = ((count_caution + count_max_caution) / tweet_count) * 100,
    prop_anger = (count_anger / tweet_count) * 100,
    prop_fear = (count_fear / tweet_count) * 100,
    prop_right = (count_right / tweet_count) * 100,
    prop_left = (count_left / tweet_count)*100,
    prop_bias = ((count_right + count_left) / tweet_count)*100
  )
```


```{r}
proportions <- c("prop_news", "prop_misinfo", "prop_anger", "prop_fear", "prop_right", "prop_left")
summary(daily_summary[, proportions])
```

```{r, fig.width=6, fig.height=6}
pdf(file.path(dst, "vars_histograms.pdf"), width = 8, height = 6)
par(mfrow = c(2,3), 
    cex.main = 1.5,    
    cex.lab = 1.3,    
    cex.axis = 1.2) 

for (p in proportions) {
  var_mean <- mean(daily_summary[[p]], na.rm = TRUE)
  
  hist(daily_summary[[p]], 
       main = paste(""),
       xlab = p,
       ylab = "",
       col = "lightblue",
       border = "black")
  abline(v = var_mean, col = "red", lwd = 2)
}
dev.off()

par(mfrow = c(1, 1))
```

## Daily time series plots for supplementary materials
```{r}
p_count <- ggplot(daily_summary, aes(x = day, y = tweet_count)) +
  geom_line() +
  custom_plot()+
  labs(tag="a", y="# tweets")+
  theme(plot.tag = element_text(size = 14, face = "bold"))


p_news <- ggplot(daily_summary, aes(x = day, y = count_ng_domains)) +
  geom_line() +
  custom_plot()+
  labs(tag="b", y="# news links")+
  theme(plot.tag = element_text(size = 14, face = "bold"))

desc_daily <- (p_count / p_news)

ggsave(file.path(dst, "desc_daily.pdf"), 
          desc_daily, 
          width = 9, height = 4.5)
```


```{r}
p_score <- ggplot(daily_summary, aes(x = day, y = mean_score)) +
  geom_line() +
  custom_plot()+
  labs(tag="a", y="NewsGuard\n score")+
  theme(plot.tag = element_text(size = 14, face = "bold"))

p_untrust <- ggplot(daily_summary, aes(x = day, y = prop_misinfo)) +
  geom_line() +
  custom_plot()+
  labs(tag="b", y="untrustworthy\n links (%)")+
  theme(plot.tag = element_text(size = 14, face = "bold"))

trustworthiness_daily <- (p_score / p_untrust)

ggsave(file.path(dst, "trustworthiness_daily.pdf"), 
          trustworthiness_daily, 
          width = 9, height = 4.5)
```

```{r}
p_anger <- ggplot(daily_summary, 
            aes(x = day, y = prop_anger)) +
  geom_line() +
  custom_plot()+
  labs(tag="a", y="angry tweets (%)")+
  theme(plot.tag = element_text(size = 14, face = "bold"))

p_fear<- ggplot(daily_summary, 
                aes(x = day, y = prop_fear)) +
  geom_line() +
  custom_plot()+
  labs(tag="b", y="fearful tweets (%)")+
  theme(plot.tag = element_text(size = 14, face = "bold"))

emotions_daily <- (p_anger / p_fear)

ggsave(file.path(dst, "emotions_daily.pdf"), 
          emotions_daily, 
          width = 9, height = 3)
```

## Weekly averages
```{r}
# summarize per week to visualize
daily_summary <- daily_summary%>%
  mutate(
    count_trustworthy = count_exceptions + count_generally_credible + count_high_credibility,
    count_untrustworthy = count_max_caution + count_caution
  )

weekly_summary <- daily_summary %>%
dplyr:::select(day, tweet_count, count_ng_domains, 
               count_trustworthy, count_untrustworthy,
               mean_score, mean_anger, mean_fear) %>%
  mutate(week = floor_date(day, "week")) %>%  # Create week column
  group_by(week) %>%
  summarise(
    across(c(tweet_count, count_ng_domains, count_trustworthy, count_untrustworthy), sum, na.rm = TRUE),
    across(c(mean_score, mean_anger, mean_fear), mean, na.rm = TRUE),
    .groups = 'drop'
  )

```

```{r}
domestic <- "gray42"
global <- "black"

events <- data.frame(
  date = as.Date(c("2019-05-17", 
                   "2020-03-12", 
                   "2020-11-02", 
                   "2021-01-06", 
                   "2021-10-09", 
                   "2022-02-24", 
                   "2022-10-21")),
  event = c("Ibiza scandal",  
            "COVID-19 lockdown",
            "Vienna attack",
            "Capitol storming", 
            "Advertisement scandal",
            "Russian invasion\n of Ukraine",
            "Woman, Life,\nFreedom protests"),
  color = c(domestic, domestic, domestic, global, domestic, global, global),
  linetype = c("dashed", "dashed", "dashed", "dotted", "dashed", "dotted", "dotted")
)

```

```{r}
max_tweets <- max(weekly_summary$tweet_count, na.rm = TRUE)

tweets_plot <- ggplot(data=weekly_summary, 
                      aes(x = week, 
                          y = tweet_count)) +
  geom_line(linewidth = 0.9) +
  geom_vline(data=events, 
             aes(xintercept = date), 
             color = events$color, 
             linetype = events$linetype,
             alpha = 0.9) +
  ggrepel:::geom_text_repel(
                  data=events, 
                  aes(x = date-50, 
                      y = max_tweets * 1.02, 
                      label = event),
                  color = events$color, 
                  size = 2.88,
                  hjust=0, 
                  direction = "y", 
                  nudge_x=0,
                  lineheight=0.7,
                  force=3, 
                  ylim = c(max_tweets * 0.95, 
                           max_tweets * 1.45),  
                  segment.color = "transparent") + 
  coord_cartesian(clip = "off") +  
  scale_y_continuous(expand = c(0, 0)) + 
  labs(tag="a", x = "", y = "# tweets") +
  custom_plot() +
  theme(
    plot.margin = margin(t = 50, r = 5, b = 0, l = 5, unit = "pt"),
    plot.tag = element_text(size = 12, face = "bold")
  )

news_plot <- ggplot(data = weekly_summary, 
                  aes(x = week, 
                      y = count_ng_domains)) +
  geom_line(linewidth = 0.9) +
  geom_vline(data = events, 
             aes(xintercept = date), 
             color = events$color, 
             linetype = events$linetype,
             alpha = 0.9) +
  labs(
    tag="b",
    x = "",
    y = "# news shared"
  ) +
  custom_plot()+
  theme(plot.tag = element_text(size = 12, face = "bold"))

score_plot <- ggplot(data = weekly_summary, 
                     aes(x = week, y = mean_score)) +
  geom_line(linewidth = 0.9) +
  geom_vline(data = events, 
             aes(xintercept = date), 
             color = events$color, 
             linetype = events$linetype, alpha = 0.9) +
  labs(
    tag="c",
    x = "time in weeks",
    y = "avg.\n trustworthiness"
  ) +
  custom_plot()+
  theme(plot.tag = element_text(size = 12, face = "bold"))

desc_plot <- (tweets_plot / news_plot / score_plot)
```

```{r}
ggsave(file.path(dst, "desc_plot.pdf"), 
          desc_plot, 
          width = 9, height = 6)
```

## Figure 1: stacked time series plots of events and trustworthy vs untrustworthy news shared per week

```{r}
plot_data <- weekly_summary %>%
  mutate(
    prop_untrustworthy = (count_untrustworthy / count_ng_domains) * 100,
    prop_trustworthy = (count_trustworthy / count_ng_domains) * 100
    ) %>%
  dplyr:::select(week, prop_trustworthy, prop_untrustworthy) %>%
  pivot_longer(
    cols = c(prop_trustworthy, prop_untrustworthy),
    names_to = "type",
    values_to = "proportion",
    names_prefix = "prop_"
  )

prop_plot <- ggplot(plot_data, aes(x = week, y = proportion, fill = type)) +
  geom_col(position = "stack", width = 7, alpha=0.8) + 
  scale_fill_manual(
    values = c("trustworthy" = "#4daf4a", "untrustworthy" = "#984ea3"),
    labels = c("trustworthy", "untrustworthy")
  ) +
  labs(
    x = "date",
    y = "trustworthiness\n of news (%)",
    tag="b") +
  custom_plot()+
  theme(
    plot.margin = margin(t = 2, r = 5, b = 5, l = 5, unit = "pt"),
    plot.tag = element_text(size = 12, face = "bold"),
    legend.position = c(0.65, 1.1), 
    legend.direction = "horizontal",
    legend.title=element_blank(),
    legend.text=element_text(size=9),
    legend.background = element_rect(fill = "transparent", color = NA), 
    legend.key.size = unit(0.25, "cm") 
  )

desc_prop_plot <- (tweets_plot / prop_plot)
ggsave(file.path(dst, "desc_prop_plot.pdf"), 
          desc_prop_plot, 
          width = 6, height = 3.5)
```

## Univariate ARIMA models and outlier detection

### Tweet count
```{r}
# extract dates
start_date <- min(daily_summary$day)
start_year <- year(start_date)
start_day  <- yday(start_date)

# convert to time series object
ts_count <- ts(daily_summary$tweet_count, 
               start = c(start_year, start_day), 
               frequency = 7 #--> use 7 to induce weekly seasonality 
               )

# fit the model without any specifications
arima_count <- auto.arima(ts_count, seasonal = TRUE)
summary(arima_count)
```

```{r}
checkresiduals(arima_count)
```

Ljung-Box: there is still significant autocorrelation. The ACF plot shows the autocorrelation of residuals at different lags (=how each residual is correlated with previous residuals: lag 1 with lag 2, etc.). All bars should fall within the blue dotted lines (95% confidence interval). There are some spikes at lag 14 and lag 28 that exceed the 95%CI but there does not seem to be an obvious temporal pattern/structure. 

The residuals fluctuate across the whole time period and there are some positive and negative spikes early and late in the series, suggesting outliers, events, or structural changes.
*--> check outliers!*

Now we're testing for outliers / abrupt changes in the time series with with tso() from tsoutliers, which gives us the type of outliers (compared to tsoutliers() from forecast)

```{r}
count_outliers <- tso(ts_count, 
                      types = c("AO", "LS", "TC"), # all types
                      tsmethod = "arima",
                      cval=3.75,
                      args.tsmethod = list(order = c(3,1,5), 
                                           seasonal = list(order = c(2, 0, 0), 
                                                           period = 7))
                      ) 

summary(count_outliers)
```


```{r}
# write function to add the date based on the date index
add_outlier_dates <- function(outlier_df, start_date) {
  outlier_df %>%
    mutate(day = as.Date(start_date) + (ind - 1))
}

count_df <- count_outliers$outliers
count_df <- add_outlier_dates(count_df, start_date)
print(count_df)
write.csv(count_df, file.path(src, "Austria-Panel-Count-Outliers.csv"), row.names = FALSE)
```

### News count 

```{r}
ts_domains <- ts(daily_summary$prop_news, 
               start = c(start_year, start_day), 
               frequency = 7)

arima_domains <- auto.arima(ts_domains, seasonal=TRUE)
summary(arima_domains)
```

```{r}
checkresiduals(arima_domains)
```

```{r}
fdGPH(residuals(arima_domains))
```

```{r}
adf_test <- ur.df(ts_domains, type = "trend")
kpss_test <- ur.kpss(ts_domains, type = "tau")

summary(adf_test)   # H0: unit root (series is I(1))
summary(kpss_test)  # H0: stationary (series is I(0))
```


```{r}
# also test on deseasonalized data
decomp <- stl(ts_domains, s.window = "periodic")
deseasonalized <- ts_domains - decomp$time.series[, "seasonal"]

adf_deseas <- ur.df(deseasonalized, type = "trend")
kpss_deseas <- ur.kpss(deseasonalized, type = "tau")
summary(adf_deseas)
summary(kpss_deseas)
```

ADF test indicates the series is stationary, whereas the KPSS test says its non-stationary. 
Test fractional integration with ``fracdiff'':

```{r}
d_estimate <- fracdiff(deseasonalized, nar = 0, nma = 0)$d
cat("Fractional integration parameter d =", d_estimate, "\n")
```


```{r}
adf_diff <- ur.df(diff(ts_domains), type = "drift")
summary(adf_diff)
```


```{r}
domain_outliers <- tso(ts_domains,
                      types = c("AO", "LS", "TC"),
                      tsmethod = "arima",
                      cval=3.75, # lower to detect COVID spike 
                      )

domains_df <- domain_outliers$outliers
domains_df <- add_outlier_dates(domains_df, start_date)
print(domains_df)
```


### Fraction of untrustworthy tweets

```{r}
ts_score <- ts(daily_summary$prop_misinfo, 
               start = c(start_year, start_day), 
               frequency = 7)

arima_score <- auto.arima(ts_score, seasonal = TRUE)
summary(arima_score)
```

```{r}
checkresiduals(arima_score)
```

*--> check for level shifts / more permanent changes*


```{r}
score_outliers <- tso(ts_score, # wouldnt converge so had to add some specifications
                      types = c("AO", "LS", "TC"),
                      tsmethod = "arima",
                      args.tsmethod = list(order = c(0,1,2),
                                           seasonal = list(order = c(2,0,0),
                                                           period = 7))
                      )

score_df <- score_outliers$outliers
score_df <- add_outlier_dates(score_df, start_date)
print(score_df)
```


```{r}
plot(score_outliers)
```


```{r}
score_intervention_index <- 1278
n_obs <- length(ts_score)

step <- ifelse(seq_len(n_obs) >= score_intervention_index, 1, 0)

ramp <- c(rep(0, score_intervention_index - 1), 
          seq(1, n_obs - score_intervention_index + 1))

xreg <- cbind(step = step, ramp = ramp)

# rerun model with dummy variable
score_intervention <- auto.arima(ts_score, 
                                 seasonal = TRUE,
                                 xreg = xreg,
                                 stepwise = FALSE) 
summary(score_intervention)
```

*step* = average value of ts_score dropped by 2.9 points 
*ramp* = after that drop, it began increasing by ~0.0002 per day

```{r}
score_intervention_index2 <- 1336
step1 <- ifelse(seq_len(n_obs) >= score_intervention_index, 1, 0)  
step2 <- ifelse(seq_len(n_obs) >= score_intervention_index2, 1, 0)  # second level shift 
xreg2 <- cbind(step1 = step1, step2 = step2)
score_2interventions <- auto.arima(ts_score, seasonal=TRUE, xreg=xreg2, stepwise=FALSE)
summary(score_2interventions)
```

```{r}
# create time series for pre-intervention
ts_score_pre <- ts(ts_score[1:(score_intervention_index - 1)],
                   frequency = frequency(ts_score),
                   start = start(ts_score))

model_cf <- auto.arima(ts_score_pre)



# forecast post-intervention period
h <- length(ts_score) - score_intervention_index + 1
fc <- forecast(model_cf, h = h)
start_fc <- time(ts_score)[score_intervention_index]
fc_ts <- ts(as.numeric(fc$mean),
            start = start_fc,
            frequency = frequency(ts_score))


# plot everything together
fitted_vals <- fitted(score_intervention)

pdf(file.path(dst, "score_intervention.pdf"), width=9, height=4.5)
plot(ts_score, 
     col = "black", 
     lwd = 1, 
     ylab = "Score", 
     xlab = "Time")

lines(fitted_vals, 
      col = "blue", 
      lwd = 2)

lines(fc_ts, 
      col = "red", 
      lwd = 2, 
      lty = "dashed")

abline(v = time(ts_score)[score_intervention_index], 
       col = "red", lty = "dashed")
abline(v = time(ts_score)[score_intervention_index2],  # add this
       col = "red", lty = "dashed", lwd = 1.5)

# legend
legend("topleft",
       legend = c("Observed", "Fitted", "Counterfactual"),
       col = c("black", "blue", "red"),
       lty = c(1, 1, 2),
       lwd = 2)

dev.off()
```


```{r}
cumulative_effect <- coef(score_2interventions)["step1"] + 
                     coef(score_2interventions)["step2"]
print(cumulative_effect); mean(ts_score_pre)
```


### Anger

```{r}
ts_anger <- ts(daily_summary$prop_anger, 
               start = c(start_year, start_day), 
               frequency = 7)

arima_anger <- auto.arima(ts_anger, seasonal=TRUE)
summary(arima_anger)
```


```{r}
checkresiduals(arima_anger)
```

```{r}
# with tso() from tsoutliers
anger_outliers <- tso(ts_anger, 
                      types = c("AO", "LS", "TC"),
                      tsmethod = "arima",
                      args.tsmethod = list(order = c(4,1,4),
                                           seasonal = list(order = c(1,0,1),
                                                           period = 7))
                      )

anger_df <- anger_outliers$outliers
anger_df <- add_outlier_dates(anger_df, start_date)
print(anger_df)
```


### Fear

```{r}
# convert to time series object
ts_fear <- ts(daily_summary$prop_fear, 
               start = c(start_year, start_day), 
               frequency = 7)

arima_fear <- auto.arima(ts_fear, seasonal=TRUE)
summary(arima_fear)
```


```{r}
checkresiduals(arima_fear)
```


```{r}
fear_outliers <- tso(ts_fear, 
                     types = c("AO", "LS", "TC"),
                     tsmethod = "arima",
                    args.tsmethod = list(order = c(1,1,3),
                                           seasonal = list(order = c(0,0,2),
                                                           period = 7))
                     ) 

fear_df <- fear_outliers$outliers
fear_df <- add_outlier_dates(fear_df, start_date)
print(fear_df)
```


## Save tables
```{r}
create_arima_table <- function(arima_model,variable_name) {
  coefs <- coef(arima_model)
  ci_matrix <- confint(arima_model)
  
  aic <- AIC(arima_model)
  bic <- BIC(arima_model)
  loglik <- arima_model$loglik
  
  coef_rows <- data.frame(
    Parameter = names(coefs),
    Coefficient = coefs,
    CI = paste0("[", round(ci_matrix[, 1], 3), ", ", round(ci_matrix[, 2], 3), "]"),
    Value = NA,
    stringsAsFactors = FALSE,
    row.names=NULL
  )
  
  stat_rows <- data.frame(
    Parameter = c("LogLik", "AIC", "BIC"),
    Coefficient = NA,
    CI = NA,
    Value = c(loglik, aic, bic),
    stringsAsFactors = FALSE
  )
  
  complete_table <- rbind(coef_rows, stat_rows)
  colnames(complete_table)[3] <- "95% CI"
  
  latex_table <- complete_table %>%
    kable(format = "latex",
          booktabs=TRUE,
          linesep = "",
          digits=3)
  
  out <- file.path(dst, paste0("arima_", variable_name, ".tex"))
  writeLines(as.character(latex_table), out)
  
  return(complete_table)
}

count_table <- create_arima_table(arima_count, "count")
```


```{r}
coefs <- coef(score_2interventions)
ci_matrix <- confint(score_2interventions)

aic <- AIC(score_2interventions)
bic <- BIC(score_2interventions)
loglik <- score_2interventions$loglik

coef_table <- data.frame(
  Parameter = names(coefs),
  Coefficient = coefs,
  CI = paste0("[", round(ci_matrix[, 1], 3), ", ", round(ci_matrix[, 2], 3), "]"),
  Value = NA,
  stringsAsFactors = FALSE,
  row.names = NULL
)

stats_table <- data.frame(
  Parameter = c("LogLik", "AIC", "BIC"),
  Coefficient = NA,
  CI = NA,
  Value = c(loglik, aic, bic),
  stringsAsFactors = FALSE,
  row.names = NULL
  )

complete_table <- rbind(coef_table, stats_table)
colnames(complete_table)[3] <- "95% CI"

latex_table <- complete_table %>%
    kable(format = "latex",
          booktabs=TRUE,
          linesep = "",
          digits=3)
  
out <- file.path(dst, "score_intervention.tex")
writeLines(as.character(latex_table), out)
```


```{r}
news_arima_table <- create_arima_table(arima_domains, "news")
anger_arima_table <- create_arima_table(arima_anger, "anger")
fear_arima_table <- create_arima_table(arima_fear, "fear")
```

## Run all of the above

```{r}
# run all the above code
```