# ===== SCRIPT OVERVIEW =====
# Task:   Fit and diagnose a VAR (Vector Autoregression) model on the daily time series.
#         Tests lag selection, residual autocorrelation, and Granger causality.
# Input:  data/Austria-Panel-Daily.csv
# Output: output/irf_plot.pdf

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 6, 
                      fig.height = 3,
                      dpi=300)

options(scipen=999)
set.seed(636)
```

```{r load packages}
library(readr)
library(scales)
library(tidyverse)
library(dplyr)
library(patchwork)
library(broom)
library(kableExtra)
library(forecast)
library(tseries)
#library(TRA)
library(vars)
library(lmtest)
library(tsoutliers)
library(tsDyn)  #VECM
```

```{r paths}
src <- here::here("data")
dst <- here::here("output")
```

```{r custom plots}
custom_plot <- function() {
  theme_classic() +  
    theme(
      axis.title.x = element_text(size = 14),
      axis.title.y = element_text(size = 14),
      axis.text.x  = element_text(size = 12),
      axis.text.y  = element_text(size = 12),
      legend.title = element_text(size = 13),
      legend.text  = element_text(size = 11)
    )
}
```

```{r}
src <- "../../data"
dst <- "../../output"
dir.create(dst, recursive = TRUE, showWarnings = FALSE)
```

```{r}
daily_summary$day <- as.Date(daily_summary$day, format = "%Y-%m-%d")
summary(daily_summary)
```

```{r}
daily_summary <- daily_summary %>% arrange(day)
head(daily_summary, 5)
```

```{r}
# create proportions
daily_summary <- daily_summary %>%
  mutate(
    prop_news = (count_ng_domains / tweet_count)*100,
    prop_trust = ((count_caution + count_max_caution) / count_ng_domains)*100,
    prop_anger = (count_anger / tweet_count)*100,
    prop_fear = (count_fear / tweet_count)*100
  )

head(daily_summary,2)
```

```{r}
start_date <- min(daily_summary$day)
start_year <- year(start_date)
start_day  <- yday(start_date)  # day of year for start
```

```{r}
ts_score <- ts(daily_summary$prop_trust, 
               start = c(start_year, start_day), 
               frequency = 7 
               )


ts_news <- ts(daily_summary$prop_news,
                  start = c(start_year, start_day), 
                  frequency = 7 
                 )

range(ts_score); range(ts_news)
```

```{r}
ts_fear <- ts(daily_summary$prop_fear,
                  start = c(start_year, start_day), 
                  frequency = 7 
                 )


ts_anger <- ts(daily_summary$prop_anger,
                  start = c(start_year, start_day), 
                  frequency = 7 
                 )


range(ts_fear); range(ts_anger)
```

```{r}
data_ts <- cbind(ts_news, ts_score, ts_fear, ts_anger)
colnames(data_ts) <- c("news", "misinfo", "fear", "anger")
plot(data_ts, main = "All time series")
```

## Relationships between variables: VAR & VECM

```{r}
# Table S8: Correlation matrix
cor_matrix <- round(cor(data_ts), 3)
print(cor_matrix)

cor_latex <- kable(as.data.frame(cor_matrix),
                   format = "latex",
                   booktabs = TRUE,
                   digits = 3)
writeLines(as.character(cor_latex), file.path(dst, "correlation_table.tex"))
```

There might be some overlap between the variables so that we move on to multivariate modeling techniques, treating all variables to be endogeneous (as a complete system if you will). 
For a strategy called Vector Autoregressive Modeling, the time series have to be stationary. Depending on the package, the data has to be transformed. When using the vars library, its function VAR allows you to include trend and seasonality components.

### Stationarity tests

```{r}
unit_root_table <- data.frame(
  Variable = character(),
  ADF = character(),
  PP = character(),
  KPSS = character(),
  stringsAsFactors = FALSE
)

variables <- list(
  "News" = ts_news,
  "Misinformation" = ts_score,
  "Fear" = ts_fear,
  "Anger" = ts_anger
)

# Run tests for each variable
for(var_name in names(variables)) {
  
  cat("Testing:", var_name, "\n")
  ts_var <- variables[[var_name]]
  
  # ADF test
  adf_result <- adf.test(ts_var)
  adf_stat <- round(adf_result$statistic, 2)
  adf_p <- ifelse(adf_result$p.value < 0.001, "<0.001", 
                  sprintf("%.3f", adf_result$p.value))
  adf_str <- paste0(adf_stat, " (", adf_p, ")")
  
  # Phillips-Perron test
  pp_result <- pp.test(ts_var)
  pp_stat <- round(pp_result$statistic, 2)
  pp_p <- ifelse(pp_result$p.value < 0.001, "<0.001", 
                  sprintf("%.3f", pp_result$p.value))
  pp_str <- paste0(pp_stat, " (", pp_p, ")")
  
  # KPSS test
  kpss_result <- kpss.test(ts_var, null = "Level")
  kpss_stat <- round(kpss_result$statistic, 2)
  kpss_p <- ifelse(kpss_result$p.value > 0.10, ">0.10",
                   ifelse(kpss_result$p.value < 0.001, "<0.001",
                          sprintf("%.3f", kpss_result$p.value)))
  kpss_str <- paste0(kpss_stat, " (", kpss_p, ")")
  
  unit_root_table <- rbind(unit_root_table, data.frame(
    Variable = var_name,
    ADF = adf_str,
    PP = pp_str,
    KPSS = kpss_str,
    stringsAsFactors = FALSE
  ))
}

unit_root_latex_table <- kable(unit_root_table, 
      format = "latex",
      booktabs = TRUE,
      align = c("l", "c", "c", "c"),
      col.names = c("Variable", "ADF", "PP", "KPSS"))

# save table
# Table S9: Unit root tests
out <- file.path(dst, "unit_root_table.tex")
writeLines(as.character(unit_root_latex_table), out)
```

All variables are non-stationary except for misinformation, which has two structural breaks. 
We test for co-integration. 


### Tests for co-integration

```{r}
n <- length(ts_score)
step1 <- ifelse(seq_len(n) >= 1278, 1, 0)  
step2 <- ifelse(seq_len(n) >= 1336, 1, 0)  
data_I1 <- cbind(ts_score, ts_fear, ts_anger)  # I(1) variables only
exog_vars <- cbind(ts_news, step1, step2)      # I(0) + break dummies

VARselect(diff(data_I1), exogen = diff(exog_vars))
```



The [forecast tutorial](https://otexts.com/fpp2/VAR.html) says: "VARselect() for selecting the number of lags (p) using four different information criteria: AIC, HQ, SC and FPE. We have met the AIC before, and SC is simply another name for the BIC (SC stands for Schwarz Criterion, after Gideon Schwarz who proposed it). HQ is the Hannan-Quinn criterion, and FPE is the “Final Prediction Error” criterion.27 Care should be taken when using the AIC as it tends to choose large numbers of lags. Instead, for VAR models, we prefer to use the BIC."

**--> Start with SC and use a simpler lag structure if in doubt as time series tend to explode with too many lags**

```{r}
summary(ca.jo(data_I1, type = "trace", ecdet = "trend",    
                 K = 6, spec = "transitory",
                 dumvar = cbind(step1, step2))  # structural break dummies
)
```

From Lütkepohl (2005, Section 6.2): "If the data have no obvious deterministic trend, or if trends are captured by other means (e.g., dummy variables for structural breaks), the constant-only specification is preferable."

```{r}
jo_results <- ca.jo(data_I1, type = "trace", ecdet = "const",    
                 K = 6, spec = "transitory",
                 dumvar = cbind(step1, step2))  # structural break dummies
        

test_stats <- jo_results@teststat
crit_values <- jo_results@cval
r_values <- 0:(length(test_stats) - 1)

jo_results_table <- data.frame(
  r = paste0("r <= ", r_values),
  Statistic = round(test_stats, 2),
  "10%" = crit_values[, 1],
  "5%" = crit_values[, 2],
  "1%" = crit_values[, 3],
  check.names = FALSE
)

jo_results_latex <- kable(jo_results_table, 
      format = "latex",
      booktabs = TRUE,
      align = c("l", "r", "r", "r", "r"))

# Table S10: Johansen cointegration test
out <- file.path(dst, "jo_results_table.tex")
writeLines(as.character(jo_results_latex), out)
```

Both specifications say r = 3 --> Use VAR in levels / this is called "full rank" co-integration.
From Lütkepohl (2005): "If the number of cointegrating vectors equals the number of variables (r = K), the system is stationary and a standard VAR in levels is appropriate."

### Finding the rightt VAR model in levels

```{r}
VARselect(data_I1, exogen = cbind(step1, step2, ts_news))
```

```{r}
summary_df <- data.frame(
  p = integer(),
  stable = logical(),
  max_root = numeric(),
  serial_chisq = numeric(),
  serial_p = numeric(),
  arch_chisq = numeric(), 
  arch_p = numeric(),
  norm_stat = numeric(),
  norm_p = character(),
  aic = numeric(),
  bic = numeric()
)

for(p in 7:9) {
  var_model <- VAR(data_I1, 
                   p = p, 
                   type = "const",
                   exogen = cbind(step1, step2, ts_news))
  
   # stability
  roots_var <- roots(var_model)
  is_stable <- all(roots_var < 1)
  max_root <- max(roots_var)
  
  # serial correlation
  serial_test <- serial.test(var_model, lags.pt = 16, type = "PT.asymptotic")
  serial_chisq <- serial_test$serial$statistic
  serial_p_value <- serial_test$serial$p.value
  serial_p <- ifelse(serial_p_value < 0.001, "<0.001", sprintf("%.3f", serial_p_value))
  
  # arch test
  arch_test <- arch.test(var_model, lags.multi = 12)
  arch_chisq <- arch_test$arch.mul$statistic
  arch_p_value <- arch_test$arch.mul$p.value
  arch_p <- ifelse(arch_p_value < 0.001, "<0.001", sprintf("%.3f", arch_p_value))
  
  # normality
  norm_test <- normality.test(var_model, multivariate.only = TRUE)
  norm_stat <- norm_test$jb.mul$JB$statistic
  norm_p_value <- norm_test$jb.mul$JB$p.value
  norm_p <- ifelse(norm_p_value < 0.001, "<0.001", sprintf("%.3f", norm_p_value))
  
  # information criteria
  aic_val <- AIC(var_model)
  bic_val <- BIC(var_model)
  
  summary_df <- rbind(summary_df, data.frame(
    p = p,
    stable = is_stable,
    max_root = max_root,
    serial_chisq = serial_chisq,
    serial_p = serial_p,
    arch_chisq = arch_chisq, 
    arch_p = arch_p,
    norm_stat = norm_stat,
    norm_p = norm_p,
    aic = aic_val,
    bic = bic_val
  ))
}

formatted_diagnostics <- data.frame(
  Lags = summary_df$p,
  Stable = ifelse(summary_df$stable, "Yes", "No"),
  Max_Root = sprintf("%.3f", summary_df$max_root),
  Serial_Test = sprintf("%.2f (%s)", summary_df$serial_chisq, summary_df$serial_p),
  ARCH_Test = sprintf("%.2f (%s)", summary_df$arch_chisq, summary_df$arch_p),
  Normality = sprintf("%.2f (%s)", summary_df$norm_stat, summary_df$norm_p),
  AIC = sprintf("%.2f", summary_df$aic),
  BIC = sprintf("%.2f", summary_df$bic)
)

diagnostics_latex <- kable(formatted_diagnostics, 
      format = "latex",
      booktabs = TRUE,
      col.names = c("Lags (p)", "Stable", "Max Root", 
                    "Portmanteau test", 
                    "ARCH test", "Normality", "AIC", "BIC"),
      align = c("c", "c", "r", "r", "r", "r", "r"),
      escape = FALSE)

# Table S11: VAR diagnostics
out <- file.path(dst, "diagnostics_table.tex")
writeLines(as.character(diagnostics_latex), out)

formatted_diagnostics
```


```{r}
final_var_model <- VAR(data_I1, 
                   p = 9, 
                   type = "const",
                   exogen = cbind(step1, step2, ts_news))

var_summary <- data.frame(
  Equation = character(),
  R_squared = numeric(),
  Adj_R_squared = numeric(),
  Residual_SE = numeric(),
  F_statistic = numeric(),
  F_p_value = numeric()
)

summary_obj <- summary(final_var_model)

for(i in 1:3) {
  eq_name <- names(summary_obj$varresult)[i]
  eq_summary <- summary_obj$varresult[[i]]
  
  var_summary <- rbind(var_summary, data.frame(
    Equation = eq_name,
    R_squared = eq_summary$r.squared,
    Adj_R_squared = eq_summary$adj.r.squared,
    Residual_SE = eq_summary$sigma,
    F_statistic = eq_summary$fstatistic[1],
    F_p_value = pf(eq_summary$fstatistic[1], 
                   eq_summary$fstatistic[2], 
                   eq_summary$fstatistic[3], 
                   lower.tail = FALSE)
  ))
}


var_summary$Equation <- c("Misinformation", "Fear", "Anger")

var_latex <- kable(var_summary,
      format = "latex",
      booktabs = TRUE,
      digits = 3,
      row.names=FALSE,
      col.names = c("Equation", "$R^2$", "Adj. $R^2$", 
                    "Residual SE", "F-statistic", "p-value"),
      escape = FALSE)

# Table S12: VAR model summary
out <- file.path(dst, "var_summary_table.tex")
writeLines(as.character(var_latex), out)

var_summary
```




## Granger causality test

```{r}
gc_misinfo <- causality(final_var_model, cause = "ts_score")
gc_fear <- causality(final_var_model, cause = "ts_fear")
gc_anger <- causality(final_var_model, cause = "ts_anger")

granger_table <- data.frame(
  Cause = c("Misinformation", "Fear", "Anger"),
  F_statistic = c(
    gc_misinfo$Granger$statistic,
    gc_fear$Granger$statistic,
    gc_anger$Granger$statistic
  ),
  df1 = c(
    gc_misinfo$Granger$parameter[1],
    gc_fear$Granger$parameter[1],
    gc_anger$Granger$parameter[1]
  ),
  df2 = c(
    gc_misinfo$Granger$parameter[2],
    gc_fear$Granger$parameter[2],
    gc_anger$Granger$parameter[2]
  ),
  p_value = c(
    gc_misinfo$Granger$p.value,
    gc_fear$Granger$p.value,
    gc_anger$Granger$p.value
  )
)

granger_formatted <- data.frame(
  Hypothesis = c(
    "Misinformation $\\rightarrow$ Fear, Anger",
    "Fear $\\rightarrow$ Misinformation, Anger",
    "Anger $\\rightarrow$ Misinformation, Fear"
  ),
  F_stat = sprintf("%.2f", granger_table$F_statistic),
  df = sprintf("(%d, %d)", granger_table$df1, granger_table$df2),
  p_value = ifelse(granger_table$p_value < 0.001, "$<$0.001",
                   sprintf("%.3f", granger_table$p_value)),
  stringsAsFactors = FALSE
)

granger_latex <- kable(granger_formatted,
      format = "latex",
      booktabs = TRUE,
      row.names = FALSE,
      align = c("l", "r", "c", "r", "l"),
      col.names = c("Null Hypothesis ($H_0$: No Granger Causality)", 
                    "F", "df", "p-value"),
      escape = FALSE) 

# Table S13: Granger causality tests
out <- file.path(dst, "granger_table.tex")
writeLines(as.character(granger_latex), out)

print(granger_table)
```

All variables are interconnected and every predictor is significant but anger is the strongest in Granger-causing others. 

Misinformation has lagged effects only.
Fear and Anger move together (same-day effects), while misinformation operates with a lag (no same-day effect).

## Impulse response function 
--> bootstrapping might take a while. 

```{r}
# Does misinfo affect fear and anger?
irf_misinfo <- irf(final_var_model, 
                   impulse = "ts_score", 
                   n.ahead = 14,
                   ortho = FALSE,
                   boot = TRUE, 
                   runs = 10000)
plot(irf_misinfo)
```

```{r}
# Does fear affect anger?
irf_fear <- irf(final_var_model, impulse = "ts_fear", 
                n.ahead = 14, boot = TRUE, 
                ortho=FALSE,
                runs = 10000)
plot(irf_fear)
```

```{r}
# Does anger feed back to misinfo?
irf_anger <- irf(final_var_model, 
                 impulse = "ts_anger", 
                 n.ahead = 14, boot = TRUE, 
                 ortho=FALSE,
                 runs = 10000)
plot(irf_anger)
```

```{r}
extract_irf_data <- function(irf_object, shock_name, response_var, shock_label, scale_factor) {
  data.frame(
    days = 0:(nrow(irf_object$irf[[shock_name]]) - 1),
    response = irf_object$irf[[shock_name]][, response_var] * scale_factor,
    lower = irf_object$Lower[[shock_name]][, response_var] * scale_factor,
    upper = irf_object$Upper[[shock_name]][, response_var] * scale_factor,
    shock = shock_label,
    response_var = response_var
  )
}


anger_data <- extract_irf_data(irf_anger, "ts_anger", "ts_score", "shock of anger", 10)
fear_data <- extract_irf_data(irf_fear, "ts_fear", "ts_score", "shock of fear", 10)
score_data <- extract_irf_data(irf_misinfo, "ts_score", "ts_score", "shock of misinfo", 10)
score_response <- bind_rows(anger_data, fear_data, score_data)

anger_data <- extract_irf_data(irf_anger, "ts_anger", "ts_anger", "shock of anger", 10)
fear_data <- extract_irf_data(irf_fear, "ts_fear", "ts_anger", "shock of fear", 10)
score_data <- extract_irf_data(irf_misinfo, "ts_score", "ts_anger", "shock of misinfo", 10)
anger_response <- bind_rows(anger_data, fear_data, score_data)

anger_data <- extract_irf_data(irf_anger, "ts_anger", "ts_fear", "shock of anger", 10)
fear_data <- extract_irf_data(irf_fear, "ts_fear", "ts_fear", "shock of fear", 10)
score_data <- extract_irf_data(irf_misinfo, "ts_score", "ts_fear", "shock of misinfo", 10)
fear_response <- bind_rows(anger_data, fear_data, score_data)
```


```{r}
score_irf <- ggplot(score_response, 
                    aes(x = days, y = response)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.3) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "gray70", alpha = 0.3) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ shock, ncol = 3, scales = "free_y") + 
  labs(x = NULL, 
       y = "misinfo (%)") +
  theme_classic() +
  theme(
    strip.background = element_rect(color = "black", fill = "gray95"),
    strip.text = element_text(size = 10, hjust = 0.5),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    axis.title.y = element_text(size = 10, margin = margin(r = 5)),
    axis.text = element_text(size = 10),
    plot.margin = margin(5, 5, 0, 5)
  )

anger_irf <- ggplot(anger_response, aes(x = days, y = response)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.3) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "gray70", alpha = 0.3) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ shock, ncol = 3, scales = "free_y") +
  labs(x = NULL,
       y = "anger (%)") +
  theme_classic() +
  theme(
    strip.background = element_blank(),  
    strip.text = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    axis.title.y = element_text(size = 10, margin = margin(r = 5)),
    axis.text = element_text(size = 10),
    plot.margin = margin(0, 5, 0, 5)
  )

fear_irf <- ggplot(fear_response, aes(x = days, y = response)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.3) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "gray70", alpha = 0.3) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ shock, ncol = 3, scales = "free_y") +
  labs(x = "days after 10% shock", 
       y = "response in fear (%)") +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    axis.title = element_text(size = 10),
    axis.title.y = element_text(margin = margin(r = 5)),
    axis.text = element_text(size = 10),
    plot.margin = margin(0, 5, 5, 5)
  )


irf_plot <- score_irf / anger_irf / fear_irf

ggsave(file.path(dst, "irf_plot.pdf"),
       plot = irf_plot,
       width = 6.5, height = 6.5)
```

### Cumulative IRF -> in 3b-boot_irf.R


### Variance Decomposition
```{r}
fevd_results <- fevd(final_var_model, n.ahead = 20)

# Create table manually
fevd_data <- data.frame(
  Variable = rep(c("Misinformation", "Fear", "Anger"), each = 5),
  Horizon = c(1, 3, 5, 7, 14),
  Own = c(
    # Misinformation at horizons 1, 5, 10, 20
    fevd_results$ts_score[c(1, 3, 5, 7, 14), "ts_score"] * 100,
    # Fear at horizons 1, 5, 10, 20
    fevd_results$ts_fear[c(1, 3, 5, 7, 14), "ts_fear"] * 100,
    # Anger at horizons 1, 5, 10, 20
    fevd_results$ts_anger[c(1, 3, 5, 7, 14), "ts_anger"] * 100
  ),
  Misinformation = c(
    # For Misinformation rows: NA (own variable)
    rep(NA, 5),
    # For Fear rows: contribution from misinformation
    fevd_results$ts_fear[c(1, 3, 5, 7, 14), "ts_score"] * 100,
    # For Anger rows: contribution from misinformation
    fevd_results$ts_anger[c(1, 3, 5, 7, 14), "ts_score"] * 100
  ),
  Fear = c(
    # For Misinformation rows: contribution from fear
    fevd_results$ts_score[c(1, 3, 5, 7, 14), "ts_fear"] * 100,
    # For Fear rows: NA (own variable)
    rep(NA, 5),
    # For Anger rows: contribution from fear
    fevd_results$ts_anger[c(1, 3, 5, 7, 14), "ts_fear"] * 100
  ),
  Anger = c(
    # For Misinformation rows: contribution from anger
    fevd_results$ts_score[c(1, 3, 5, 7, 14), "ts_anger"] * 100,
    # For Fear rows: contribution from anger
    fevd_results$ts_fear[c(1, 3, 5, 7, 14), "ts_anger"] * 100,
    # For Anger rows: NA (own variable)
    rep(NA, 5)
  )
)

# Format
fevd_data$Own <- sprintf("%.1f", fevd_data$Own)
fevd_data$Misinformation <- ifelse(is.na(fevd_data$Misinformation), 
                                   "—", 
                                   sprintf("%.1f", fevd_data$Misinformation))
fevd_data$Fear <- ifelse(is.na(fevd_data$Fear), 
                         "—", 
                         sprintf("%.1f", fevd_data$Fear))
fevd_data$Anger <- ifelse(is.na(fevd_data$Anger), 
                          "—", 
                          sprintf("%.1f", fevd_data$Anger))


fevd_latex <- kable(fevd_data,
      format = "latex",
      booktabs = TRUE,
      row.names = FALSE,
      align = c("l", "c", "c", "c", "c"),
      col.names = c("Variable", "Horizon", "Own",
                    "Misinformation", "Fear", "Anger"),
      escape = FALSE)

# Table S14: Variance decomposition (FEVD)
out <- file.path(dst, "fevd_table.tex")
writeLines(as.character(fevd_latex), out)
```

## Run all of the above

```{r}
# run all the above code
```