# ===== SCRIPT OVERVIEW =====
# Task:   Visualise user-level emotion dynamics features and bootstrap regression results.
#         Shows feature distributions, correlation matrix, PCA, and model coefficients.
# Input:  data/Austria-Panel-Users-Features-Tweet.csv
#         ./user_models_tweet_boot.rds  (output of 6-boot_features_models.R)
# Output: ./users/features_hist.png
#         ./users/corr_matrix.png
#         ./users/outcome_bar.png
#         ./users/user_sharing_proportions.png
#         ./users/pca_scree_plot.png
#         ./users/boot_results_plot.png
#         output/features_hist.pdf
#         output/pca_scree_plot.pdf

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, message = FALSE,
                      fig.width = 6.5, 
                      fig.height = 4.5,
                      dpi=300)
```

```{r packages}
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(corrplot)
library(car)
library(FactoMineR)
library(factoextra)
library(boot)
library(pROC)
library(kableExtra)
````

```{r paths}
src <- "../../data"
dst <- "../../output"
file <- "Austria-Panel-Users-Features-Tweet.csv"
dir.create("./users", showWarnings = FALSE, recursive = TRUE)

features <- read_csv(file.path(src, file),
                     col_types = cols(
                       .default = col_double()
                     )) 
```


```{r}
custom_plot <- function() {
  theme_classic() +  
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x  = element_text(size = 11),
      axis.text.y  = element_text(size = 11),
      legend.title = element_text(size = 12),
      legend.text  = element_text(size = 11)
    )
}
```

# ===== DATA OVERVIEW =====
```{r}
cat("Total users:", nrow(features), "\n\n")

cat("Feature summary:\n")
print(summary(features))

cat("\nMissing values:\n")
print(colSums(is.na(features)))
```


# ===== DISTRIBUTION PLOTS =====
```{r}
corr_features <- features %>%
  dplyr:::select(-author_id)

dist_data <- corr_features %>%
  pivot_longer(everything(), names_to = "feature", values_to = "value")

# remove avg_untrustworthy and avg_biased
pred_data <- dist_data %>%
  filter(feature != "n_news_shared" & feature != "n_untrustworthy_shared" & feature != "n_biased_shared")

dist_plot <- ggplot(pred_data, aes(x = value)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.9) +
  facet_wrap(~ feature, scales = "free", ncol = 4) +
  labs(x = "value",
       y = "frequency") +
  custom_plot() +
  theme(
     axis.text.x  = element_text(size = 8),
     axis.text.y = element_text(size = 8)
        )

print(dist_plot)
```



# ===== SUMMARY STATISTICS TABLE =====
```{r}
summary_stats <- corr_features %>%
  summarise(across(everything(), 
                   list(mean = ~mean(., na.rm = TRUE),
                        sd = ~sd(., na.rm = TRUE),
                        median = ~median(., na.rm = TRUE),
                        min = ~min(., na.rm = TRUE),
                        max = ~max(., na.rm = TRUE),
                        n_missing = ~sum(is.na(.)))))

print(summary_stats)
```


# ===== REGRESSION ANALYSIS =====
```{r}
features$shares_news <- as.integer(features$n_news_shared > 0)
features$shares_untrustworthy <- as.integer(features$n_untrustworthy_shared > 0)
features$shares_biased <- as.integer(features$n_biased_shared > 0)

table(features$shares_news)
table(features$shares_untrustworthy)
table(features$shares_biased)
```

```{r}
total_users <- nrow(features)

shares_news_count <- sum(features$shares_news)
shares_news_pct <- round(shares_news_count / total_users * 100, 1)

shares_untrustworthy_count <- sum(features$shares_untrustworthy)
shares_untrustworthy_pct <- round(shares_untrustworthy_count / total_users * 100, 1)

shares_biased_count <- sum(features$shares_biased)
shares_biased_pct <- round(shares_biased_count / total_users * 100, 1)

user_table <- data.frame(
  `News type` = c("Any news source", 
                  "Untrustworthy sources (NewsGuard < 60)", 
                  "Partisan sources (left/right bias)"),
  `Users (n)` = c(shares_news_count, 
                              shares_untrustworthy_count, 
                              shares_biased_count),
  `Users (\\%)` = c(shares_news_pct, 
                              shares_untrustworthy_pct, 
                              shares_biased_pct),
  check.names = FALSE
)

print(user_table)
```


# ===== MULTICOLLINEARITY =====
```{r}
m_news <- glm(shares_news ~ anger_baseline + 
                            fear_baseline +
                            anger_variability +
                            fear_variability +
                            anger_instability +
                            fear_instability +
                            anger_inertia +
                            fear_inertia,
                  data = features,
                  family = binomial(link = "logit"))

vif(m_news)
```

```{r}
emotion_vars <- c('anger_baseline', 'anger_variability', 
                  'anger_instability', 'anger_inertia',
                  'fear_baseline', 'fear_variability',
                  'fear_instability', 'fear_inertia')

pca_data <- features[complete.cases(features[, emotion_vars]), emotion_vars]

pca_result <- prcomp(pca_data, 
                     scale. = TRUE, 
                     center = TRUE)

summary(pca_result)
```
```{r}
pca_scree <- fviz_eig(pca_result, addlabels = TRUE,
 ylim = c(0, 61))+
  custom_plot()+
  theme(plot.title=element_blank())

```
```{r}
print(pca_result$rotation)
```

```{r}
variance_explained <- summary(pca_result)$importance
print(variance_explained[, 1:4])
```

# ===== MODEL EVALUATION =====

```{r}
complete_features <- features[complete.cases(features[, 
                        c('shares_untrustworthy',  # shouldnt have NaNs but just in case
                          'shares_news',
                          'shares_biased',
                          'anger_baseline', 'fear_baseline', # could have NaN if not enough data points
                          'anger_variability', 'fear_variability',
                          'anger_instability', 'fear_instability',
                          'anger_inertia', 'fear_inertia')]), ]

z_features <- complete_features %>%
  mutate(across(c(anger_baseline, fear_baseline, 
                  anger_variability, fear_variability,
                  anger_instability, fear_instability,
                  anger_inertia, fear_inertia),
                ~as.numeric(scale(.))))

df <- z_features

# whats the number of users? 
nrow(df)
```
```{r}
m_news_base <- glm(shares_news ~ anger_baseline + fear_baseline,
               data = df, family = binomial)

m_news_variability <- glm(shares_news ~ anger_variability + fear_variability,
               data = df, family = binomial)

m_news_instability <- glm(shares_news ~ anger_instability + fear_instability,
               data = df, family = binomial)

m_news_inertia <- glm(shares_news ~ anger_inertia + fear_inertia,
               data = df, family = binomial)

vif(m_news_base); vif(m_news_variability);vif(m_news_instability); vif(m_news_inertia)
```
```{r}
m_misinfo_base <- glm(shares_untrustworthy ~ anger_baseline + fear_baseline,
               data = df, family = binomial)

m_misinfo_variability <- glm(shares_untrustworthy ~ anger_variability + fear_variability,
               data = df, family = binomial)

m_misinfo_instability <- glm(shares_untrustworthy ~ anger_instability + fear_instability,
               data = df, family = binomial)

m_misinfo_inertia <- glm(shares_untrustworthy ~ anger_inertia + fear_inertia,
               data = df, family = binomial)

vif(m_misinfo_base); vif(m_misinfo_variability);vif(m_misinfo_instability); vif(m_misinfo_inertia)
```

```{r}
m_partisan_base <- glm(shares_biased ~ anger_baseline + fear_baseline,
               data = df, family = binomial)

m_partisan_variability <- glm(shares_biased ~ anger_variability + fear_variability,
               data = df, family = binomial)

m_partisan_instability <- glm(shares_biased ~ anger_instability + fear_instability,
               data = df, family = binomial)

m_partisan_inertia <- glm(shares_biased ~ anger_inertia + fear_inertia,
               data = df, family = binomial)

vif(m_partisan_base); vif(m_partisan_variability);vif(m_partisan_instability); vif(m_partisan_inertia)
```



# ===== ADD BOOTSTRAP UNCERTAINTY =====
```{r}
boot_objects <- readRDS("users/user_models_tweet_boot.rds")

boot_news_base <- boot_objects$boot_news_base
boot_news_var <- boot_objects$boot_news_var
boot_news_inst <- boot_objects$boot_news_inst
boot_news_inertia <- boot_objects$boot_news_inertia
boot_misinfo_base <- boot_objects$boot_misinfo_base
boot_misinfo_var <- boot_objects$boot_misinfo_var
boot_misinfo_inst <- boot_objects$boot_misinfo_inst
boot_misinfo_inertia <- boot_objects$boot_misinfo_inertia
boot_partisan_base <- boot_objects$boot_partisan_base
boot_partisan_var <- boot_objects$boot_partisan_var
boot_partisan_inst <- boot_objects$boot_partisan_inst
boot_partisan_inertia <- boot_objects$boot_partisan_inertia
```

```{r}
get_boot_cis <- function(boot_obj, outcome_name, feature_name) {
  param_names <- names(boot_obj$t0)
  param_names <- param_names[param_names != "(Intercept)"] # skip intercept
  indices <- which(names(boot_obj$t0) %in% param_names)
  
  ci_df <- data.frame(
    outcome = outcome_name,
    feature = feature_name,
    parameter = param_names,
    estimate = boot_obj$t0[indices],
    lower = apply(boot_obj$t[, indices, drop=FALSE], 2, quantile, probs = 0.025),
    upper = apply(boot_obj$t[, indices, drop=FALSE], 2, quantile, probs = 0.975)
  )
  
  rownames(ci_df) <- NULL
  return(ci_df)
}


boot_results <- bind_rows(
  get_boot_cis(boot_news_base, "any news", "baseline"),
  get_boot_cis(boot_news_var, "any news", "variability"),
  get_boot_cis(boot_news_inst, "any news", "instability"),
  get_boot_cis(boot_news_inertia, "any news", "inertia"),
  
  get_boot_cis(boot_misinfo_base, "untrustworthy news", "baseline"),
  get_boot_cis(boot_misinfo_var, "untrustworthy news", "variability"),
  get_boot_cis(boot_misinfo_inst, "untrustworthy news", "instability"),
  get_boot_cis(boot_misinfo_inertia, "untrustworthy news", "inertia"),
  
  get_boot_cis(boot_partisan_base, "partisan news", "baseline"),
  get_boot_cis(boot_partisan_var, "partisan news", "variability"),
  get_boot_cis(boot_partisan_inst, "partisan news", "instability"),
  get_boot_cis(boot_partisan_inertia, "partisan news", "inertia")
)
```


# ===== RESULTS VISUALIZATION =====
```{r}
boot_results_plot <- boot_results %>%
  mutate(
    OR = exp(estimate),
    OR_lower = exp(lower),
    OR_upper = exp(upper),
    emotion = case_when(
      grepl("anger", parameter) ~ "anger",
      grepl("fear", parameter) ~ "fear"
    ),
    feature = factor(feature, levels = c("baseline", "variability", "instability", "inertia")),
    outcome = factor(outcome, levels = c("any news", "partisan news", "untrustworthy news"))
  )

print(boot_results_plot)
```

```{r}
or_plot <- ggplot(boot_results_plot, aes(x = OR, y = feature, color = emotion)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_errorbarh(aes(xmin = OR_lower, xmax = OR_upper),
                 position = position_dodge(width = 0.5),
                 height = 0,
                 linewidth = 0.7) +
  facet_wrap(~ outcome, ncol = 3) +
  labs(x = "Odds Ratio (95% CI)",
       y = "Emotional Feature",
       color = "") +
  scale_color_manual(values = c("anger" = "#e41a1c", "fear" = "#377eb8")) +
  custom_plot() +
  theme(
    legend.position = "bottom",
    legend.box.margin = margin(t = -10, r = 0, b = 0, l = 0),
    axis.text = element_text(color="black"),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(face = "bold", size = 11))

print(or_plot)

# Figure 2a
ggsave(file.path(dst, "features_boot_results.pdf"),
       or_plot,
       width = 6.5, height = 4.5, dpi = 300)
```


## SAVE FOR PAPER
```{r}
ggsave(file.path(dst, "features_hist.pdf"), 
          dist_plot, 
          width = 6.5, height = 4,
          dpi=300)
```
```{r}
user_latex <- kable(user_table, 
                    format = "latex", 
                    booktabs = TRUE,
                    align = c("l", "c", "c"),
                    escape = FALSE)

# Table S15: News behavior of users (sharing proportions)
out <- file.path(dst, "user_table.tex")
writeLines(as.character(user_latex), out)
```

```{r}
# stargazer wont work on one of the models so we create our own function
create_latex_table <- function(models, model_names, filename) {

  # extract coefficients
  coefs <- lapply(models, function(m) {
    summ <- summary(m)$coefficients
    data.frame(
      est = summ[, "Estimate"],
      se = summ[, "Std. Error"],
      pval = summ[, "Pr(>|z|)"],
      term = rownames(summ)
    )
  })

  # add sig. stars
  get_stars <- function(p) {
    if (p < 0.001) return("***")
    if (p < 0.01) return("**")
    if (p < 0.05) return("*")
    return("")
  }

  latex <- ""
  latex <- paste0(latex, "\\begin{tabular}{l", paste(rep("c", length(models)), collapse=""), "}\n")
  latex <- paste0(latex, "\\hline\n")

  latex <- paste0(latex, " & ", paste(model_names, collapse = " & "), " \\\\\n")
  latex <- paste0(latex, "\\hline\n")


  for (term in c("anger", "fear")) {
    latex <- paste0(latex, ifelse(term == "anger", "Anger", "Fear"))
    for (i in 1:length(models)) {
      coef_row <- coefs[[i]][grep(term, coefs[[i]]$term, ignore.case = TRUE), ]
      if (nrow(coef_row) > 0) {
        stars <- get_stars(coef_row$pval)
        latex <- paste0(latex, " & ", sprintf("%.3f", coef_row$est), stars,
                        " (", sprintf("%.3f", coef_row$se), ")")
      } else {
        latex <- paste0(latex, " & ")
      }
    }
    latex <- paste0(latex, " \\\\\n")
  }

  latex <- paste0(latex, "Constant")
  for (i in 1:length(models)) {
    int_row <- coefs[[i]][coefs[[i]]$term == "(Intercept)", ]
    stars <- get_stars(int_row$pval)
    latex <- paste0(latex, " & ", sprintf("%.3f", int_row$est), stars,
                    " (", sprintf("%.3f", int_row$se), ")")
  }
  latex <- paste0(latex, " \\\\\n")

  latex <- paste0(latex, "\\hline\n")

  latex <- paste0(latex, "Observations")
  for (m in models) {
    latex <- paste0(latex, " & ", format(nobs(m), big.mark=","))
  }
  latex <- paste0(latex, " \\\\\n")

  latex <- paste0(latex, "AIC")
  for (m in models) {
    latex <- paste0(latex, " & ", sprintf("%.1f", AIC(m)))
  }
  latex <- paste0(latex, " \\\\\n")

  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "\\multicolumn{", length(models)+1, "}{l}{\\footnotesize $^{***}$p$<$0.001; $^{**}$p$<$0.01; $^{*}$p$<$0.05} \\\\\n")
  latex <- paste0(latex, "\\end{tabular}\n")

  # Tables S16 (news), S17 (untrustworthy), S18 (partisan)
  writeLines(latex, file.path(dst, filename))
  cat("LaTeX table written to", filename, "\n")
}


create_latex_table(
  list(m_news_base, m_news_variability, m_news_instability, m_news_inertia),
  c("Baseline", "Variability", "Instability", "Inertia"),
  "table_news_models.tex"
)

create_latex_table(
  list(m_misinfo_base, m_misinfo_variability, m_misinfo_instability, m_misinfo_inertia),
  c("Baseline", "Variability", "Instability", "Inertia"),
  "table_misinfo_models.tex"
)

create_latex_table(
  list(m_partisan_base, m_partisan_variability, m_partisan_instability, m_partisan_inertia),
  c("Baseline", "Variability", "Instability", "Inertia"),
  "table_partisan_models.tex"
)
```

```{r}
ggsave(file.path(dst,"pca_scree_plot.pdf"), 
       pca_scree)
```

```{r}
# run all
print("Running all analysis steps...")
```
