# ===== SCRIPT OVERVIEW =====
# Task:   Visualise event window results and compare LMER model coefficients.
#         Shows emotion distributions before/after news events; caterpillar plots;
#         supersharer CDF and individual-level trajectories.
# Input:  data/Austria-Panel-Windows-1h-10min.csv.gz
#         data/Austria-Panel-Windows-24h-10min.csv.gz
#         data/windows_10_1h/*.csv   (output of 8-boot_window_models.R)
#         data/windows_10_24h/*.csv
# Output: output/window_tweet_distributions.pdf
#         output/boot_means_comparison.pdf
#         output/emotions_models_compared.pdf
#         output/windows_emotions_coefs.pdf
#         output/windows_caterpillar_combined.pdf
#         output/individual_plot.pdf


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dpi=300,
                      fig.width = 6.5,
                      fig.height = 4.5)

set.seed(636)
````

```{r libraries}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(scales)
library(kable)
library(kableExtra)
library(lme4)
library(performance)
library(DHARMa)
library(lmerTest)
```

```{r fig}
custom_plot <- function() {
  theme_classic() +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text.x  = element_text(size = 11),
      axis.text.y  = element_text(size = 11),
      legend.title = element_text(size = 12),
      legend.text  = element_text(size = 11)
    )
}

rating_colors <- c("untrustworthy" = "#8b7ab8", 
                   "trustworthy" = "#2a9d8f")
emotion_colors <- c("anger" = "#e41a1c", 
                    "fear" = "#377eb8")
emotion_colors_caps <- c("Anger" = "#e41a1c", 
                         "Fear" = "#377eb8")
```

```{r paths}
src        <- "../../data"
dst        <- "../../output"   # figures and tables
models_dir <- "windows_models" # intermediate model objects (relative to scripts/analysis/)
```

## Data Overview

```{r data}
short_window <- read_csv(file.path(src, "Austria-Panel-Windows-1h-10min.csv.gz"),
                        col_types = cols(
                          author_id = col_character(),
                          event_id = col_character(),
                          event_score = col_double(),
                          mean_anger_before = col_double(),
                          mean_fear_before = col_double(),
                          mean_anger_after = col_double(),
                          mean_fear_after = col_double()
                        )) %>%
  mutate(
    untrustworthy = (100 - event_score) / 100
  )

cat("Number of unique users:", n_distinct(short_window$author_id), "\n")
cat("Number of unique events:", n_distinct(short_window$event_id), "\n")
```

```{r data}
long_window <- read_csv(file.path(src, "Austria-Panel-Windows-24h-10min.csv.gz"),
                        col_types = cols(
                          author_id = col_character(),
                          event_id = col_character(),
                          event_score = col_double(),
                          mean_anger_before = col_double(),
                          mean_fear_before = col_double(),
                          mean_anger_after = col_double(),
                          mean_fear_after = col_double()
                        )) %>%
  mutate(
    untrustworthy = (100 - event_score) / 100
  )

cat("Number of unique users:", n_distinct(long_window$author_id), "\n")
cat("Number of unique events:", n_distinct(long_window$event_id), "\n")
```

## Distribution of events and tweets per event in each window

```{r distributions}
prepare_long_shape <- function(data) {
  data %>%
    select(author_id, event_id,
           mean_anger_before, mean_anger_after,
           mean_fear_before, mean_fear_after) %>%
    pivot_longer(
      cols = c(mean_anger_before, mean_anger_after,
               mean_fear_before, mean_fear_after),
      names_to = c("emotion", "window"),
      names_pattern = "mean_(.*)_(before|after)",
      values_to = "value"
    ) %>%
    filter(!is.na(value)) %>%  # Remove missing values
    mutate(
      emotion = factor(emotion, levels = c("anger", "fear")),
      window = factor(window, levels = c("before", "after"))
    )
}

window_long <- prepare_long_shape(long_window)
window_short <- prepare_long_shape(short_window)

count_tweets <- function(data, window_label) {
  data %>%
    group_by(author_id, window) %>%
    summarise(
      n_events = n_distinct(event_id),
      n_tweets = n(),
      .groups = "drop"
    ) %>%
    mutate(window_size = window_label)
}

window_counts_24h <- count_tweets(window_long, "24h")
window_counts_1h <- count_tweets(window_short, "1h")
```

```{r}
p_24h <- ggplot(window_counts_24h, aes(x = n_tweets)) +
  geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
  scale_x_log10(labels = scales::comma) +
  labs(
    title = "b.",
    x = "# of tweets per user (log)",
    y = "frequency"
  ) +
  facet_wrap(~ window, scales = "free_y") +
  custom_plot()+
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
  )

p_1h <- ggplot(window_counts_1h, aes(x = n_tweets)) +
  geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
  scale_x_log10(labels = scales::comma) +
  labs(
    title = "a.",
    x = "# of tweets per user (log)",
    y = "frequency"
  ) +
  facet_wrap(~ window, scales = "free_y") +
  custom_plot()+
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
  )

p_combined <- p_1h / p_24h

print(p_combined)
```

```{r}
ggsave(file.path(dst, "./window_tweet_distributions.pdf"),
       plot = p_combined,
       width = 6.5, height = 5.5)
```

## Distribution of emotions before and after events in each window
```{r emotions}
# plot distributions of emotions before and after events for each window
p_emotions_1h <- ggplot(window_short, aes(x = value, fill = emotion)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(~ window, scales = "free_y") +
  labs(
    title = "a.",
    x = "Emotion score",
    y = "frequency"
  ) +
  scale_fill_manual(values = emotion_colors) +
  custom_plot()+
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
  )

p_emotions_24h <- ggplot(window_long, aes(x = value, fill = emotion)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(~ window, scales = "free_y") +
  labs(
    title = "b.",
    x = "Emotion score",
    y = "frequency"
  ) +
  scale_fill_manual(values = emotion_colors) +
  custom_plot()+
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
  )

p_emotions_combined <- p_emotions_1h / p_emotions_24h

print(p_emotions_combined)
```

```{r}
ggsave(file.path(dst, "window_emotions_comparison.png"),
       plot = p_emotions_combined,
       width = 6.5, height = 5.5)

```

## Sensitivity analysis of emotion means
```{r}
# define function to bootstrap means and plot across sample sizes
boot_mean_stability <- function(data, list_col, emotion, window) {

  # extract lists of values from dump column in df 
  all_values <- data %>%
    filter(lengths(.data[[list_col]]) > 0) %>%  
    select(.data[[list_col]]) %>%
    unnest(cols = all_of(list_col)) %>%
    pull() %>%
    as.numeric() %>%
    na.omit()

  bootstrap_means <- expand.grid(
    n = seq(2, 20, by = 1),
    rep = 1:1000
  ) %>%
    mutate(mean = NA_real_)
  
  sample_sizes <- unique(bootstrap_means$n)
  
  for (i in seq_along(sample_sizes)) {
    n <- sample_sizes[i]
    bootstrap_means$mean[bootstrap_means$n == n] <- replicate(1000, {
      mean(sample(all_values, n, replace = TRUE), na.rm = TRUE)
    })
  }
  
  stability_summary <- bootstrap_means %>%
    group_by(n) %>%
    summarise(
      mean_mean = mean(mean, na.rm = TRUE),
      sd_mean   = sd(mean, na.rm = TRUE),
      ci_lower  = quantile(mean, 0.025, na.rm = TRUE),
      ci_upper  = quantile(mean, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  p <- ggplot(stability_summary, aes(x = n, y = mean_mean)) +
    geom_line() +
    geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.2) +
    labs(
      title = paste(window),
      x = "Number of tweets sampled",
      y = paste(emotion, "mean ± 95% CI"),
    ) +
    custom_plot()+
    theme(
      plot.title = element_text(size = 16, face = "bold"),
      axis.title.x = element_text(size = 11),
      axis.title.y = element_text(size = 11)
    )
  
  return(p)
}
```

```{r}
res_anger_1h <- boot_mean_stability(short_window, "anger_before_list",
                                         "anger", "a.")
```

```{r}
res_fear_1h <- boot_mean_stability(short_window, 
                                    "fear_before_list", 
                                    "fear", "")
```


```{r}
res_anger_24h <- boot_mean_stability(long_window, 
                                    "anger_before_list", 
                                    "anger", "b.")
```

```{r}
res_fear_24h <- boot_mean_stability(long_window, 
                                    "fear_before_list", 
                                    "fear", "")
```

```{r}
p_means_combined <- (res_anger_1h + res_fear_1h) / (res_anger_24h + res_fear_24h)

print(p_means_combined)
```

```{r}
ggsave(file.path(dst, "boot_means_comparison.pdf"),
       plot = p_means_combined,
       width = 6.5, height = 5.5)
```

## Statistical model results

```{r before sharing}
short_coefs <- read_csv(file = file.path(models_dir, "windows_10_1h", "coefs_window_models.csv")) %>%
  dplyr:::select(emotion, window, model, estimate, ci_lower, ci_upper)%>%
  mutate(window = factor(window, levels = c("before", "after", "change")),
         emotion = factor(emotion, levels = c("anger", "fear")))

long_coefs <- read_csv(file = file.path(models_dir, "windows_10_24h", "coefs_window_models.csv")) %>%
  dplyr:::select(emotion, window, model, estimate, ci_lower, ci_upper)%>%
  mutate(window = factor(window, levels = c("before", "after", "change")),
         emotion = factor(emotion, levels = c("anger", "fear")))


p_short_models <- ggplot(short_coefs, aes(
                               x = estimate,
                               y = emotion, 
                               color = model)) +
  facet_wrap(~ window, scales = "fixed" ) +
  geom_vline(xintercept = 0, 
              linetype = "dashed",
              color = "black", 
              linewidth = 0.5) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = ci_lower, xmax = ci_upper), 
                width = 0.2, linewidth = 0.8,
                position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("lmer" = "#1b9e77", "lm" = "#d95f02")) +
  scale_y_discrete(labels = c("anger", "fear")) +
  labs(
    title = "a.",
    x = "Coefficient",
    y = "",
    color = "model type"
  ) +
  custom_plot() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 11),
    strip.text = element_text(size = 11, face = "bold")  
  )+
  theme(legend.position = "none")


p_long_models <- ggplot(long_coefs, aes(
                               x = estimate,
                               y = emotion, 
                               color = model)) +
  facet_wrap(~ window, scales = "fixed") +
  geom_vline(xintercept = 0, 
              linetype = "dashed", 
              color = "black", 
              linewidth = 0.5) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = ci_lower, xmax = ci_upper), 
                width = 0.2, linewidth = 0.8,
                position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("lmer" = "#1b9e77", "lm" = "#d95f02")) +
  scale_y_discrete(labels = c("anger", "fear")) +
  labs(
    title = "b.",
    x = "Coefficient",
    y = "",
    color = "model type"
  ) +
  custom_plot() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 11),
    strip.text = element_text(size = 11, face = "bold")  
  )+
  theme(legend.position = "bottom",
        legend.box.spacing = unit(0.1, "cm"),
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(nrow = 1, byrow = TRUE))



p_models_combined <- p_short_models / p_long_models

print(p_models_combined)
```

```{r}
ggsave(file.path(dst, "emotions_models_compared.pdf"),
       plot = p_models_combined,
       height = 4.5, width = 6.5)
```

## Model diagnostics

```{r}
# load all models
m_anger_24h_before <- readRDS(file.path(models_dir, "windows_10_24h", "lmer_anger_before.rds"))
m_fear_24h_before <- readRDS(file.path(models_dir, "windows_10_24h", "lmer_fear_before.rds"))
m_anger_1h_before <- readRDS(file.path(models_dir, "windows_10_1h", "lmer_anger_before.rds"))
m_fear_1h_before <- readRDS(file.path(models_dir, "windows_10_1h", "lmer_fear_before.rds"))
m_anger_24h_after <- readRDS(file.path(models_dir, "windows_10_24h", "lmer_anger_after.rds"))
m_fear_24h_after <- readRDS(file.path(models_dir, "windows_10_24h", "lmer_fear_after.rds"))
m_anger_1h_after <- readRDS(file.path(models_dir, "windows_10_1h", "lmer_anger_after.rds"))
m_fear_1h_after <- readRDS(file.path(models_dir, "windows_10_1h", "lmer_fear_after.rds"))
m_anger_24h_change <- readRDS(file.path(models_dir, "windows_10_24h", "lmer_anger_change.rds"))
m_fear_24h_change <- readRDS(file.path(models_dir, "windows_10_24h", "lmer_fear_change.rds"))
m_anger_1h_change <- readRDS(file.path(models_dir, "windows_10_1h", "lmer_anger_change.rds"))
m_fear_1h_change <- readRDS(file.path(models_dir, "windows_10_1h", "lmer_fear_change.rds"))
```

### Random effects structure
```{r re}
# analyze random effects structure
extract_random_effects <- function(model, emotion, window, size) {
  vc <- as.data.frame(VarCorr(model))
  between_sd <- vc$sdcor[vc$grp == "author_id"]
  within_sd <- vc$sdcor[vc$grp == "Residual"]
  between_var <- vc$vcov[vc$grp == "author_id"]
  within_var <- vc$vcov[vc$grp == "Residual"]

  icc_val <- icc(model)$ICC_adjusted

  re <- ranef(model)$author_id[,1]
  re_min <- min(re)
  re_max <- max(re)
  re_sd <- sd(re)
  
  n_obs <- nobs(model)
  n_groups <- nlevels(model@frame$author_id)
  

  data.frame(
    Emotion = emotion,
    Window = window,
    Size = size,
    N_obs = n_obs,
    N_users = n_groups,
    Between_SD = between_sd,
    Within_SD = within_sd,
    Between_Var = between_var,
    Within_Var = within_var,
    ICC = icc_val,
    RE_Min = re_min,
    RE_Max = re_max,
    RE_SD = re_sd
  )
}

model_list <- list(
  list(m_anger_24h_before, "Anger", "Before", "24h"),
  list(m_anger_1h_before, "Anger", "Before", "1h"),
  list(m_fear_24h_before, "Fear", "Before", "24h"),
  list(m_fear_1h_before, "Fear", "Before", "1h"),
  list(m_anger_24h_after, "Anger", "After", "24h"),
  list(m_anger_1h_after, "Anger", "After", "1h"),
  list(m_fear_24h_after, "Fear", "After", "24h"),
  list(m_fear_1h_after, "Fear", "After", "1h"),
  list(m_anger_24h_change, "Anger", "Change", "24h"),
  list(m_anger_1h_change, "Anger", "Change", "1h"),
  list(m_fear_24h_change, "Fear", "Change", "24h"),
  list(m_fear_1h_change, "Fear", "Change", "1h")
)

re_table <- do.call(rbind, lapply(model_list, function(x) {
  extract_random_effects(x[[1]], x[[2]], x[[3]], x[[4]])
})) %>%
  mutate(
    N_obs = format(N_obs, big.mark = ","),
    N_users = format(N_users, big.mark = ","),
    Between_SD = round(Between_SD, 3),
    Within_SD = round(Within_SD, 3),
    Between_Var = round(Between_Var, 4),
    Within_Var = round(Within_Var, 4),
    ICC = round(ICC, 3),
    RE = round(RE_Min, 3),
    RE_Max = round(RE_Max, 3),
    RE_SD = round(RE_SD, 3)
  )

print(re_table)
```

```{r}
create_re_latex <- function(re_table) {
  
  latex <- "\\begin{tabular}{lllccccccc}\n"
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "Emotion & Window & Size & ICC & Between SD & Within SD & RE Range & N obs & N users \\\\\n")
  latex <- paste0(latex, "\\hline\n")
  
  for (i in 1:nrow(re_table)) {
    row <- re_table[i,]

    latex <- paste0(latex,
                    row$Emotion, " & ",
                    row$Window, " & ",
                    row$Size, " & ",
                    sprintf("%.3f", row$ICC), " & ",
                    sprintf("%.3f", row$Between_SD), " & ",
                    sprintf("%.3f", row$Within_SD), " & ",
                    "[", sprintf("%.2f", row$RE_Min), ", ",
                    sprintf("%.2f", row$RE_Max), "] & ",
                    row$N_obs, " & ",
                    row$N_users, " \\\\\n")
  }
  
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "\\end{tabular}\n")
  
  # Table S22: Random effects structure
  writeLines(latex, file.path(dst, "windows_random_effects.tex"))
  cat("LaTeX table saved to windows_random_effects.tex\n")
  
  return(latex)
}

create_re_latex(re_table)
```


## Visualizing the fixed effects 

```{r lmer}
p_short_coefs <- short_coefs%>%
  filter(model == "lmer") %>%
  ggplot(aes(x = estimate,
             y = emotion,
             color = emotion)) +
  facet_wrap(~ window, scales = "fixed") +
  geom_vline(xintercept = 0, 
              linetype = "dashed", 
              color = "black", 
              linewidth = 0.5) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = ci_lower, xmax = ci_upper), 
                width = 0.2, linewidth = 0.7) +
  scale_color_manual(values = emotion_colors) +
  scale_y_discrete(labels = c("anger", "fear")) +
  labs(
    title = "a.",
    x = "Coefficient",
    y = ""
  ) +
  guides(color = "none") +
  custom_plot() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 11),
    strip.text = element_text(size = 11, face = "bold")  
  )

p_long_coefs <- long_coefs%>%
  filter(model == "lmer") %>%
  ggplot(aes(x = estimate,
             y = emotion,
             color = emotion)) +
  facet_wrap(~ window, scales = "fixed") +
  geom_vline(xintercept = 0, 
              linetype = "dashed", 
              color = "black", 
              linewidth = 0.5) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = ci_lower, xmax = ci_upper), 
                width = 0.2, linewidth = 0.7) +
  scale_color_manual(values = emotion_colors) +
  scale_y_discrete(labels = c("anger", "fear")) +
  labs(
    title = "b.",
    x = "Coefficient",
    y = ""
  ) +
  guides(color = "none") +
  custom_plot() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 11),
    strip.text = element_text(size = 11, face = "bold")
  )

p_coefs_combined <- p_short_coefs / p_long_coefs

print(p_coefs_combined)
```

```{r}
short_coefs$window_size <- "24h"
long_coefs$window_size <- "1h"

coefs_combined <- bind_rows(short_coefs, long_coefs) %>%
  filter(model == "lmer") %>%
  mutate(
    emotion = factor(emotion, levels = c("anger", "fear")),
    window = factor(window, levels = c("before", "after", "change")),
    window_size = factor(window_size, levels = c("1h", "24h"))
  )

pd <- position_dodge(width = 0.35)
p_coefs <- ggplot(
  coefs_combined,
  aes(
    x = estimate,
    y = emotion,
    color = emotion,
    alpha = window_size
  )
) +
  facet_wrap(~ window, scales = "free_x") +

  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "black",
    linewidth = 0.7
  ) +

  geom_point(
    size = 3,
    position = pd
  ) +

  geom_errorbarh(
    aes(xmin = ci_lower, xmax = ci_upper),
    height = 0.2,
    linewidth = 0.9,
    position = pd
  ) +

  scale_color_manual(values = emotion_colors) +
  scale_alpha_manual(values = c("24h" = 0.5, "1h" = 1)) +

  scale_y_discrete(labels = c("Anger", "Fear")) +
  labs(
    x = "Coefficient",
    y = ""
  ) +
  guides(color = "none") +
  custom_plot() +
  theme(
    axis.title.x = element_text(size = 11),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = # move to upper right in the plot area
      c(0.9, 0.9),
    legend.title=element_blank(),
    # make x-axis ticks smaller
    axis.text.x = element_text(size = 9)
  )

print(p_coefs)
```

```{r}
ggsave(file.path(dst, "windows_emotions_coefs.pdf"),
       plot = p_coefs,
       width = 6.5, height = 4.5)
```

## Creating tables summarizing the model results
```{r}
# stargazer wont work on one of the models so we create our own function
create_before_table <- function(anger_24_model, fear_24_model, anger_1_model, fear_1_model, 
                                short_coefs, long_coefs, filename) {
  anger_24_coefs <- broom.mixed::tidy(anger_24_model, effects = "fixed")
  fear_24_coefs <- broom.mixed::tidy(fear_24_model, effects = "fixed")
  anger_1_coefs <- broom.mixed::tidy(anger_1_model, effects = "fixed")
  fear_1_coefs <- broom.mixed::tidy(fear_1_model, effects = "fixed")

  # helper to look up bootstrap CIs from the coefs CSV
  get_boot_ci <- function(coefs_csv, emotion_val, window_val) {
    row <- coefs_csv %>%
      filter(emotion == emotion_val, window == window_val, model == "lmer")
    if (nrow(row) == 0) return(c(NA, NA))
    c(row$ci_lower[1], row$ci_upper[1])
  }

  format_row <- function(label, window, time, coefs, term_name, boot_ci) {
    row <- coefs[coefs$term == term_name, ]
    if (nrow(row) == 0) return("")
    est <- row$estimate[1]
    se <- row$std.error[1]
    ci_str <- paste0("[", sprintf("%.3f", boot_ci[1]), ", ", sprintf("%.3f", boot_ci[2]), "]")
    paste0(label, " & ", window, " & ", time, " & ",
           sprintf("%.3f", est), " & ", sprintf("%.3f", se), " & ", ci_str, " & ",
           sprintf("%.1f", row$df[1]), " & ", sprintf("%.3f", row$p.value[1]), " \\\\\n")
  }

  latex <- ""
  latex <- paste0(latex, "\\begin{tabular}{lccccccr}\n")
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "Emotion & Window & Size & Estimate & SE & 95\\% CI & df & p\\\\\n")
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, format_row("Anger", "Before", "1h", anger_1_coefs, "mean_anger_before",
                                    get_boot_ci(short_coefs, "anger", "before")))
  latex <- paste0(latex, format_row("Anger", "Before", "24h", anger_24_coefs, "mean_anger_before",
                                    get_boot_ci(long_coefs, "anger", "before")))
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, format_row("Fear", "Before", "1h", fear_1_coefs, "mean_fear_before",
                                    get_boot_ci(short_coefs, "fear", "before")))
  latex <- paste0(latex, format_row("Fear", "Before", "24h", fear_24_coefs, "mean_fear_before",
                                    get_boot_ci(long_coefs, "fear", "before")))
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "\\end{tabular}\n")

  writeLines(latex, file.path(dst, filename))
  cat("LaTeX table written to", file.path(dst, filename), "\n")
  return(latex)
}

# call with the coefs passed in
# Table S19: LMER predicting untrustworthy sharing from emotion before event
create_before_table(m_anger_24h_before, m_fear_24h_before, m_anger_1h_before, m_fear_1h_before,
                    short_coefs, long_coefs, "emotions_before.tex")

```

```{r}
create_after_table <- function(anger_1_model, anger_24_model, fear_1_model, fear_24_model,
                               short_coefs, long_coefs, filename) {
  anger_24_coefs <- broom.mixed::tidy(anger_24_model, effects = "fixed")
  fear_24_coefs <- broom.mixed::tidy(fear_24_model, effects = "fixed")
  anger_1_coefs <- broom.mixed::tidy(anger_1_model, effects = "fixed")
  fear_1_coefs <- broom.mixed::tidy(fear_1_model, effects = "fixed")

  get_boot_ci <- function(coefs_csv, emotion_val, window_val) {
    row <- coefs_csv %>%
      filter(emotion == emotion_val, window == window_val, model == "lmer")
    if (nrow(row) == 0) return(c(NA, NA))
    c(row$ci_lower[1], row$ci_upper[1])
  }

  format_row <- function(label, window, time, coefs, term_name, boot_ci) {
    row <- coefs[coefs$term == term_name, ]
    if (nrow(row) == 0) return("")
    est <- row$estimate[1]
    se <- row$std.error[1]
    ci_str <- paste0("[", sprintf("%.3f", boot_ci[1]), ", ", sprintf("%.3f", boot_ci[2]), "]")
    paste0(label, " & ", window, " & ", time, " & ",
           sprintf("%.3f", est), " & ", sprintf("%.3f", se), " & ", ci_str, " & ",
           sprintf("%.1f", row$df[1]), " & ", sprintf("%.3f", row$p.value[1]), " \\\\\n")
  }

  latex <- ""
  latex <- paste0(latex, "\\begin{tabular}{llccccccr}\n")
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "Emotion & Window & Size & Estimate & SE & 95\\% CI & df & p\\\\\n")
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, format_row("Anger", "After", "1h", anger_1_coefs, "untrustworthy",
                                    get_boot_ci(short_coefs, "anger", "after")))
  latex <- paste0(latex, format_row("Anger", "After", "24h", anger_24_coefs, "untrustworthy",
                                    get_boot_ci(long_coefs, "anger", "after")))
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, format_row("Fear", "After", "1h", fear_1_coefs, "untrustworthy",
                                    get_boot_ci(short_coefs, "fear", "after")))
  latex <- paste0(latex, format_row("Fear", "After", "24h", fear_24_coefs, "untrustworthy",
                                    get_boot_ci(long_coefs, "fear", "after")))
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "\\end{tabular}\n")

  writeLines(latex, file.path(dst, filename))
  cat("LaTeX table written to", file.path(dst, filename), "\n")
  return(latex)
}

# Table S20: LMER predicting emotion after sharing from trustworthiness
create_after_table(m_anger_1h_after, m_anger_24h_after, m_fear_1h_after, m_fear_24h_after,
                   short_coefs, long_coefs, "emotions_after.tex")
```

```{r}
create_change_table <- function(anger_1_model, anger_24_model, fear_1_model, fear_24_model,
                                short_coefs, long_coefs, filename) {
  anger_1_coefs <- broom.mixed::tidy(anger_1_model, effects = "fixed")
  anger_24_coefs <- broom.mixed::tidy(anger_24_model, effects = "fixed")
  fear_1_coefs <- broom.mixed::tidy(fear_1_model, effects = "fixed")
  fear_24_coefs <- broom.mixed::tidy(fear_24_model, effects = "fixed")

  get_boot_ci <- function(coefs_csv, emotion_val, window_val) {
    row <- coefs_csv %>%
      filter(emotion == emotion_val, window == window_val, model == "lmer")
    if (nrow(row) == 0) return(c(NA, NA))
    c(row$ci_lower[1], row$ci_upper[1])
  }

  format_row <- function(label, window, time, coefs, term_name, boot_ci) {
    row <- coefs[coefs$term == term_name, ]
    if (nrow(row) == 0) return("")
    est <- row$estimate[1]
    se <- row$std.error[1]
    ci_str <- paste0("[", sprintf("%.3f", boot_ci[1]), ", ", sprintf("%.3f", boot_ci[2]), "]")
    paste0(label, " & ", window, " & ", time, " & ",
           sprintf("%.3f", est), " & ", sprintf("%.3f", se), " & ", ci_str, " & ",
           sprintf("%.1f", row$df[1]), " & ", sprintf("%.3f", row$p.value[1]), " \\\\\n")
  }

  latex <- ""
  latex <- paste0(latex, "\\begin{tabular}{lccccccr}\n")
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "Emotion & Window & Size & Estimate & SE & 95\\% CI & df & p\\\\\n")
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, format_row("Anger", "Change", "1h", anger_1_coefs, "untrustworthy",
                                    get_boot_ci(short_coefs, "anger", "change")))
  latex <- paste0(latex, format_row("Anger", "Change", "24h", anger_24_coefs, "untrustworthy",
                                    get_boot_ci(long_coefs, "anger", "change")))
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, format_row("Fear", "Change", "1h", fear_1_coefs, "untrustworthy",
                                    get_boot_ci(short_coefs, "fear", "change")))
  latex <- paste0(latex, format_row("Fear", "Change", "24h", fear_24_coefs, "untrustworthy",
                                    get_boot_ci(long_coefs, "fear", "change")))
  latex <- paste0(latex, "\\hline\n")
  latex <- paste0(latex, "\\end{tabular}\n")

  writeLines(latex, file.path(dst, filename))
  cat("LaTeX table written to", file.path(dst, filename), "\n")

  return(latex)
}

# Table S21: LMER testing within-person change in emotion around sharing events
create_change_table(m_anger_1h_change, m_anger_24h_change, m_fear_1h_change, m_fear_24h_change,
                    short_coefs, long_coefs, "emotions_change.tex")
```



## Create caterpillar plots with random effects for main text 

```{r}
# define function to extract bootstrapped random effects
process_bootstrap_re <- function(boot_file, original_model, model_name) {
  
  cat("Processing", model_name, "...\n")
  boot_results <- readRDS(boot_file)
  random_boot <- lapply(boot_results, function(x) x$random_effects)
  re_orig <- ranef(original_model)$author_id
  fixed_int <- fixef(original_model)[1]
  original_users <- rownames(re_orig)
  n_users <- length(original_users)
  n_boot <- length(random_boot)
  
  cat("  ", n_users, "users,", n_boot, "bootstrap iterations\n")
  
  # create matrix to store bootstrap samples
  # rows = bootstrap iterations, cols = users
  boot_matrix <- matrix(NA, nrow = n_boot, ncol = n_users)
  colnames(boot_matrix) <- original_users
  
  for (i in 1:n_boot) {
    re_boot <- random_boot[[i]]
    if (!is.null(re_boot)) {
      matched_idx <- match(original_users, names(re_boot)) # match to orig users
      boot_matrix[i, ] <- re_boot[matched_idx]
    }
  }
  
  boot_cis <- t(apply(boot_matrix, 2, quantile, 
                      probs = c(0.025, 0.975), na.rm = TRUE))
  
  re_df <- data.frame(
    author_id = original_users,
    predicted_baseline = fixed_int + re_orig[,1],
    ci_lower = boot_cis[,1],
    ci_upper = boot_cis[,2],
    model = model_name,
    stringsAsFactors = FALSE
  )
  return(re_df)
}
```


```{r}
anger_before_re <- process_bootstrap_re(
  boot_file = file.path(models_dir, "windows_10_1h", "boot_lmer_anger_before.rds"),
  original_model = m_anger_1h_before,
  model_name = "anger_before"
)

fear_before_re <- process_bootstrap_re(
  boot_file = file.path(models_dir, "windows_10_1h", "boot_lmer_fear_before.rds"),
  original_model = m_fear_1h_before,
  model_name = "fear_before"
)

anger_after_re <- process_bootstrap_re(
  boot_file = file.path(models_dir, "windows_10_1h", "boot_lmer_anger_after.rds"),
  original_model = m_anger_1h_after,
  model_name = "anger_after"
)

fear_after_re <- process_bootstrap_re(
  boot_file = file.path(models_dir, "windows_10_1h", "boot_lmer_fear_after.rds"),
  original_model = m_fear_1h_after,
  model_name = "fear_after"
)
```
```{r}
# define function to plot the caterpillar plot per model 
plot_caterpillar <- function(re_df, title,
                             n_display = NULL) {
  
  re_plot <- re_df %>%
    arrange(predicted_baseline)
  
  # take a sample bc its too many users to plot
  re_plot <- re_plot %>%
    mutate(decile = ntile(predicted_baseline, 10)) %>% # sample across deciles of baseline
    group_by(decile) %>%
    slice_sample(n = ceiling(n_display / 10)) %>%
    ungroup() %>%
    arrange(predicted_baseline)
  
  re_plot$rank <- 1:nrow(re_plot)
  
  # population average (red line)
  pop_avg <- mean(re_plot$predicted_baseline, na.rm = TRUE)
  
  p <- ggplot(re_plot, aes(x = rank, y = predicted_baseline)) +
    geom_hline(yintercept = pop_avg, 
               linetype = "dashed", color = "red", linewidth = 0.5) +
    geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                    size = 0.01, alpha = 0.6, linewidth = 0.6,
                    color = "black") +
    labs(
      x = "users (ranked by baseline)",
      y = "predicted\nbaseline",
      title = title
    ) +
    custom_plot() +
    theme(
      plot.title = element_text(size = 11),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.y = element_text(size = 11, margin = margin(r = 3))
    )

  return(p)
}
```


```{r}
p1 <- plot_caterpillar(
  anger_before_re,
  title = "untrustworthy by anger",
  n_display = 1000
) + labs(tag = "a")

p2 <- plot_caterpillar(
  anger_after_re,
  title = "anger by untrustworthy",
  n_display = 1000
) + labs(tag = "b")

p3 <- plot_caterpillar(
  fear_before_re,
  title = "untrustworthy by fear",
  n_display = 1000
)+
  theme(
    axis.title.y = element_blank()
  ) + labs(tag = "c")
p4 <- plot_caterpillar(
  fear_after_re,
  title = "fear by untrustworthy",
  n_display = 1000
) +
  labs(tag = "d") +
  theme(
    axis.title.y = element_blank()
  )

# arrange them all in 2x2 grid
p_caterpillar <- (p1 | p2) / (p3 | p4) &
  theme(plot.tag = element_text(size = 16, face = "bold"))

print(p_caterpillar)
```

```{r}
ggsave(file.path(dst, "windows_caterpillar_combined.pdf"),
       plot = p_caterpillar,
       width = 6.5, height = 3.5)
```

## Create supersharer panels (a, b)

```{r}
users_daily <- read_csv(file.path(src, "Austria-Panel-Users-Daily.csv.gz"),
                        col_types = cols(
                          author_id = col_character(),
                          .default = col_double()
                        ))

users_summary <- users_daily %>%
  group_by(author_id) %>%
  summarise(
    total_misinfo_count = sum(n_untrustworthy, na.rm = TRUE),
    total_trustworthy_count = sum(n_trustworthy, na.rm = TRUE),
    .groups = "drop"
  )

untrustworthy_by_user <- users_summary %>%
  select(author_id, total_misinfo_count) %>%
  arrange(desc(total_misinfo_count)) %>%
  mutate(
    cumsum_count = cumsum(total_misinfo_count),
    cumsum_prop = cumsum_count / sum(total_misinfo_count),
    user_rank = row_number(),
    percentile = user_rank / n() * 100,
    source_type = "untrustworthy"
  )

threshold_untrustworthy <- min(which(untrustworthy_by_user$cumsum_prop >= 0.80))
percentile_untrustworthy <- untrustworthy_by_user$percentile[threshold_untrustworthy]

trustworthy_by_user <- users_summary %>%
  select(author_id, total_trustworthy_count) %>%
  filter(total_trustworthy_count > 0) %>%
  arrange(desc(total_trustworthy_count)) %>%
  mutate(
    cumsum_count = cumsum(total_trustworthy_count),
    cumsum_prop = cumsum_count / sum(total_trustworthy_count),
    user_rank = row_number(),
    percentile = user_rank / n() * 100,
    source_type = "trustworthy"
  )

threshold_trustworthy <- min(which(trustworthy_by_user$cumsum_prop >= 0.80))
percentile_trustworthy <- trustworthy_by_user$percentile[threshold_trustworthy]

users_combined <- bind_rows(untrustworthy_by_user, trustworthy_by_user)
````

```{r}
p_supersharers_cdf <- ggplot(users_combined,
                             aes(x = percentile,
                                 y = cumsum_prop,
                                 color = source_type)) +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = 0.8,
             linetype = "dashed",
             color = "black") +
  scale_color_manual(values = c("trustworthy" = "#2a9d8f",
                                "untrustworthy" = "#8b7ab8")) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    x = "users (%)",
    y = "news shared (%)",
    color = NULL,
    tag = "b"
  ) +
  annotate("text",
           x = 15, y = 0.84,
           label = paste0(round(percentile_untrustworthy, 2), "% vs ",
                          round(percentile_trustworthy, 2), "% of users"),
           size = 3.5,
           hjust = 0) +
  custom_plot() +
  theme(
    legend.direction = "horizontal",
    legend.position = c(0.5, 1.1),
    plot.tag = element_text(size = 16, face = "bold")
  )

print(p_supersharers_cdf)
```

```{r}
supersharers <- read_csv(file.path(src, "Austria-Panel-Supersharers.csv"),
                         col_types = cols(
                           author_id = col_character(),
                           group = col_character(),
                           .default = col_double()
                         ))

anger_plot_data <- supersharers %>%
  select(author_id, group,
         anger_baseline, anger_variability, anger_instability) %>%
  pivot_longer(cols = c(anger_baseline, anger_variability, anger_instability),
               names_to = "metric",
               values_to = "value") %>%
  mutate(
    group = tolower(group),
    metric = gsub("anger_", "", metric),
    metric = factor(metric, levels = c("baseline", "instability", "variability")),
    metric_label = as.character(metric)
  ) %>%
  filter(is.finite(value))

anger_stats <- anger_plot_data %>%
  group_by(group, metric_label) %>%
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop")

anger_wilcox <- anger_plot_data %>%
  group_by(metric_label) %>%
  nest() %>%
  mutate(
    wilcox_test = map(data, ~{
      trustworthy <- .x %>% filter(group == "trustworthy") %>% pull(value)
      untrustworthy <- .x %>% filter(group == "untrustworthy") %>% pull(value)
      wilcox.test(trustworthy, untrustworthy)
    }),
    p_value = map_dbl(wilcox_test, ~.x$p.value),
    significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  ) %>%
  select(metric_label, p_value, significance)

anger_density_peaks <- anger_plot_data %>%
  group_by(metric_label) %>%
  do({
    d <- density(.$value, na.rm = TRUE)
    data.frame(peak_y = max(d$y))
  }) %>%
  ungroup()

anger_stars <- anger_wilcox %>%
  left_join(
    anger_stats %>%
      pivot_wider(names_from = group, values_from = mean_value),
    by = "metric_label"
  ) %>%
  left_join(anger_density_peaks, by = "metric_label") %>%
  mutate(
    x = pmax(trustworthy, untrustworthy) + 0.01,
    y = if_else(metric_label == "baseline", peak_y * 0.9, peak_y * 0.99),
    label = significance
  )

group_colors <- c("untrustworthy" = "#8b7ab8", "trustworthy" = "#2a9d8f")

p_supersharers_anger <- ggplot(anger_plot_data, aes(x = value, fill = group, color = group)) +
  geom_density(alpha = 0.35, linewidth = 0.5) +
  geom_vline(data = anger_stats,
             aes(xintercept = mean_value, color = group),
             linetype = "dashed", linewidth = 0.8) +
  geom_text(data = anger_stars,
            aes(x = x, y = y, label = label),
            size = 3, hjust = 0, vjust = 0.5,
            inherit.aes = FALSE) +
  facet_wrap(~ metric_label, ncol = 1, scales = "free_y") +
  scale_fill_manual(values = group_colors) +
  scale_color_manual(values = group_colors) +
  labs(
    x = "anger",
    y = "density",
    y = NULL,
    fill = NULL,
    color = NULL,
    tag = "c"
  ) +
  custom_plot() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size = 9),
    panel.grid = element_blank(),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(size = 10),
    legend.position = "none",
    plot.tag = element_text(size = 16, face = "bold")
  )

print(p_supersharers_anger)
```


```{r}
# add the features plot as panel a
boot_objects <- readRDS("users/user_models_tweet_boot.rds")

boot_news_base <- boot_objects$boot_news_base
boot_news_var <- boot_objects$boot_news_var
boot_news_inst <- boot_objects$boot_news_inst
boot_news_inertia <- boot_objects$boot_news_inertia
boot_misinfo_base <- boot_objects$boot_misinfo_base
boot_misinfo_var <- boot_objects$boot_misinfo_var
boot_misinfo_inst <- boot_objects$boot_misinfo_inst
boot_misinfo_inertia <- boot_objects$boot_misinfo_inertia
boot_partisan_base <- boot_objects$boot_partisan_base
boot_partisan_var <- boot_objects$boot_partisan_var
boot_partisan_inst <- boot_objects$boot_partisan_inst
boot_partisan_inertia <- boot_objects$boot_partisan_inertia

get_boot_cis <- function(boot_obj, outcome_name, feature_name) {
  param_names <- names(boot_obj$t0)
  param_names <- param_names[param_names != "(Intercept)"] # skip intercept
  indices <- which(names(boot_obj$t0) %in% param_names)
  
  ci_df <- data.frame(
    outcome = outcome_name,
    feature = feature_name,
    parameter = param_names,
    estimate = boot_obj$t0[indices],
    lower = apply(boot_obj$t[, indices, drop=FALSE], 2, quantile, probs = 0.025),
    upper = apply(boot_obj$t[, indices, drop=FALSE], 2, quantile, probs = 0.975)
  )
  
  rownames(ci_df) <- NULL
  return(ci_df)
}

boot_results <- bind_rows(
  get_boot_cis(boot_news_base, "any news", "baseline"),
  get_boot_cis(boot_news_var, "any news", "variability"),
  get_boot_cis(boot_news_inst, "any news", "instability"),
  get_boot_cis(boot_news_inertia, "any news", "inertia"),
  
  get_boot_cis(boot_misinfo_base, "untrustworthy news", "baseline"),
  get_boot_cis(boot_misinfo_var, "untrustworthy news", "variability"),
  get_boot_cis(boot_misinfo_inst, "untrustworthy news", "instability"),
  get_boot_cis(boot_misinfo_inertia, "untrustworthy news", "inertia"),
  
  get_boot_cis(boot_partisan_base, "partisan news", "baseline"),
  get_boot_cis(boot_partisan_var, "partisan news", "variability"),
  get_boot_cis(boot_partisan_inst, "partisan news", "instability"),
  get_boot_cis(boot_partisan_inertia, "partisan news", "inertia")
)

boot_results_plot <- boot_results %>%
  mutate(
    OR = exp(estimate),
    OR_lower = exp(lower),
    OR_upper = exp(upper),
    emotion = case_when(
      grepl("anger", parameter) ~ "anger",
      grepl("fear", parameter) ~ "fear"
    ),
    feature = factor(feature, levels = c("baseline", "variability", "instability", "inertia")),
    outcome = factor(outcome, levels = c("any news", "partisan news", "untrustworthy news"))
  )
```

```{r}
or_plot <- ggplot(boot_results_plot, aes(x = OR, y = feature, color = emotion)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_errorbar(aes(xmin = OR_lower, xmax = OR_upper),
                position = position_dodge(width = 0.5),
                width = 0, 
                linewidth = 0.7) +
  facet_wrap(~ outcome, ncol = 3) +
  labs(x = "Odds Ratio (95% CI)",
       y = "",
       color = "") +
  scale_color_manual(values = c("anger" = "#e41a1c", 
                                "fear" = "#377eb8")) +
  scale_y_discrete(limits = c("inertia", 
                              "instability", 
                              "variability", 
                              "baseline")) +
  custom_plot() +
  theme(
    legend.position = "top", 
    legend.box.margin = margin(t = -10, r = 0, b = -10, l = 0),
    legend.text = element_text(size = 10),
    axis.text = element_text(color="black"),
    strip.background = element_rect(fill = "white", color = "black"),
    strip.text = element_text(
      #face = "bold", 
      size = 11))
```

```{r}
bottom_row <- (p_supersharers_cdf | p_supersharers_anger) +
  plot_layout(widths = c(1.2, 0.8))

combined_fig2 <- ((or_plot + labs(tag = "a") +
    theme(plot.tag = element_text(size = 16, face = "bold"))) /
  bottom_row) +
  plot_layout(heights = c(1, 0.9))

print(combined_fig2)
```
```{r}
ggsave(file.path(dst, "individual_plot.pdf"),
       plot = combined_fig2)
```

## Run all chunks above
```{r}
# run the whole script to generate all tables and figures
```
